# R and R Studio {#randrstudio}

## What is R?

### R (programming language), [Wikipedia](https://en.wikipedia.org/wiki/R_(programming_language))

* **R is a programming language** and **free software** environment for **statistical computing and graphics** supported by the R Foundation for Statistical Computing. 

* The R language is widely used among statisticians and data miners for developing statistical software and data analysis.

* A **GNU package**, the official R software environment is written primarily in C, Fortran, and R itself (thus, it is partially self-hosting) and is freely available under the GNU General Public License. 



### History of R and more

"R Programming for Data Science" by Roger Peng

* [Chapter 2. History and Overview of R](https://bookdown.org/rdpeng/rprogdatascience/history-and-overview-of-r.html)
* [Overview and History of R: Youtube video](https://www.youtube.com/watch?v=STihTnVSZnI&feature=youtu.be)



## Why R? -- Responses by Hadley Wickham

### [r4ds](https://r4ds.had.co.nz/introduction.html#python-julia-and-friends): R is a great place to start your data science journey because

* R is an environment designed from the ground up to support data science. 
* R is not just a programming language, but it is also an interactive environment for doing data science. 
* To support interaction, R is a much more flexible language than many of its peers. 



### Why R today?

When you talk about choosing programming languages, I always say you shouldn’t pick them based on technical merits, but rather pick them based on the community. And I think the R community is like really, really strong, vibrant, free, welcoming, and embraces a wide range of domains. So, if there are like people like you using R, then your life is going to be much easier. That’s the first reason. 

**Interview**: ["Advice to Young (and Old) Programmers, H. Wickham"](https://www.r-bloggers.com/2018/08/advice-to-young-and-old-programmers-a-conversation-with-hadley-wickham/)



## What is RStudio? https://posit.com

> RStudio is an integrated development environment, or IDE, for R programming. 

### R Studio (Wikipedia)

RStudio is an integrated development environment (IDE) for R, a programming language for statistical computing and graphics. It is available in two formats: RStudio Desktop is a regular desktop application while RStudio Server runs on a remote server and allows accessing RStudio using a web browser.



## Installation of R and R Studio

### R Installation

To download R, go to CRAN, the comprehensive R archive network. CRAN is composed of a set of mirror servers distributed around the world and is used to distribute R and R packages. Don’t try and pick a mirror that’s close to you: instead use the cloud mirror, https://cloud.r-project.org, which automatically figures it out for you.

A new major version of R comes out once a year, and there are 2-3 minor releases each year. It’s a good idea to update regularly.

### R Studio Installation

Download and install it from http://www.rstudio.com/download. 

RStudio is updated a couple of times a year. When a new version is available, RStudio will let you know.



## R Studio

### The First Step
1. Start R Studio Application
2. Top Menu: File > New Project > New Directory > New Project > _Directory name or Browse the directory and choose the parent directory you want to create the directory_

### When You Start the Project
1. Go to the directory you created
2. Double click _'Directory Name'.Rproj
  
Or,

1. Start R Studio
2. File > Open Project (or choose from Recent Project)

_In this way the working directory of the session is set to the project directory and R can search releted files without difficulty_ (`getwd()`, `setwd()`)



## Posit Cloud

RStudio Cloud is a lightweight, cloud-based solution that allows anyone to do, share, teach and learn data science online.

### Cloud Free

* Up to 15 projects total
* 1 shared space (5 members and 10 projects max)
* 15 project hours per month
* Up to 1 GB RAM per project
* Up to 1 CPU per project
* Up to 1 hour background execution time



### How to Start Posit Cloud

1. Go to https://posit.cloud/
2. Sign Up: _top right_
  - Email address or Google account
3. New Project: _Project Name_
4. R Console


## Let's Get Started 

Start RStudio and create a project, or login to Posit Cloud and create a project.



### The First Examples

Input the following codes into Console in the left bottom pane.

* The first two:

```{r}
head(cars)
```



```{r}
str(cars)
```



* Two more:

```{r}
summary(cars)
```



```{r eval=FALSE}
plot(cars)
```

```{r cars_plot, echo=FALSE}
plot(cars)
```



* And three more:

```{r eval=FALSE}
plot(cars) # cars: Speed and Stopping Distances of Cars
abline(lm(cars$dist~cars$speed))
```
```{r echo=FALSE}
plot(cars) # cars: Speed and Stopping Distances of Cars
abline(lm(cars$dist~cars$speed))
```



```{r}
lm(cars$dist~cars$speed)
```



```{r}
summary(lm(cars$dist~cars$speed))
```



#### Brief Explanation

* `cars`: A pre-installed data which is a part of package `datasets`.

* `head(cars)`: The first six rows of the pre-installed data `cars`.
* `str(cars)`: The pre-installed data `cars` data structure.
* `summary(cars)`: The summary of the pre-installed data `cars`.
* `plot(cars)`: A scatter plot of the pre-installed data `cars`.
  - `plot(cars$dist~cars$speed)`
  - `cars$dist`, `cars$[[2]]`, `cars[,2]` are same
* `abline(lm(cars$dist~cars$speed))`: Add a regression line of a linear model
* `lm(cars$dist~cars$speed)`: The equation of the regression line
* `summary(lm(cars$dist~cars$speed)`: The summary of the linear regression model

#### Histograms

`cars` is a data frame consisting of two columns, `dist` and `speed`.

```{r, eval=FALSE}
hist(cars$dist)
```
```{r, echo=FALSE}
hist(cars$dist)
```



```{r, eval=FALSE}
hist(cars$speed)
```
```{r, echo=FALSE}
hist(cars$speed)
```



#### View and help

* `View(cars)`
* `?cars`: same as `help(cars)`
* `??cars`: same as `help.search("cars")

#### `datasets`

`datasets` is a pre-installed datasets which contains a variety of datasets.

* `?datasets`
* `library(help = "datasets")`

* `data()` shows all data already attached and available.

By `library(help = "datasets")` and/or `data()`, you can see the complete list of data in `datasets`.

### Practicum

Pick a data in the datasets package and try

* `head()`
* `str()`
* `summary()`

and some more.



### `iris`

```{r}
head(iris)
```



```{r}
str(iris)
```



```{r}
summary(iris)
```



Can you plot?

```{r eval=FALSE}
plot(iris$Sepal.Length, iris$Sepal.Width)
```
```{r echo=FALSE}
plot(iris$Sepal.Length, iris$Sepal.Width)
```

