#  `dplyr` {#dplyr}

> The `dplyr` is a package to transform data. It can combine data as well. We will treat the second feature late in Chapter \@ref(tidyr). The package `dplyr` is a part of the `tidyverse` packages, and you do not need to install it separately.

```{r}
library(tidyverse)
```


## `dplyr` [Overview](https://dplyr.tidyverse.org)

dplyr is a grammar of data manipulation, providing a consistent set of verbs that help you solve the most common data manipulation challenges:

* `select()` picks variables based on their names.
* `filter()` picks cases based on their values.
* `mutate()` adds new variables that are functions of existing variables
* `summarise()` reduces multiple values down to a single summary.
* `arrange()` changes the ordering of the rows.
* `group_by()` takes an existing tbl and converts it into a grouped tbl.

You can learn more about them in vignette("dplyr"). As well as these single-table verbs, dplyr also provides a variety of two-table verbs, which you can learn about in vignette("two-table").

If you are new to dplyr, the best place to start is [the data transformation chapter in R for data science](http://r4ds.had.co.nz/transform.html).


## [`select`](https://dplyr.tidyverse.org/reference/select.html): Subset columns using their names and types

Helper Function	| Use	| Example
---|-------|--------
-	| Columns except	| select(babynames, -prop)
:	| Columns between (inclusive)	| select(babynames, year:n)
contains() |	Columns that contains a string |	select(babynames, contains("n"))
ends_with()	| Columns that ends with a string	| select(babynames, ends_with("n"))
matches()	| Columns that matches a regex |	select(babynames, matches("n"))
num_range()	| Columns with a numerical suffix in the range | Not applicable with babynames
one_of() |	Columns whose name appear in the given set |	select(babynames, one_of(c("sex", "gender")))
starts_with()	| Columns that starts with a string	| select(babynames, starts_with("n"))



## [`filter`](https://dplyr.tidyverse.org/reference/filter.html): Subset rows using column values

Logical operator	| tests	| Example
--|-----|---
>	| Is x greater than y? |	x > y
>=	| Is x greater than or equal to y? |	x >= y
<	| Is x less than y?	| x < y
<=	| Is x less than or equal to y? | 	x <= y
==	| Is x equal to y? |	x == y
!=	| Is x not equal to y? |	x != y
is.na()	| Is x an NA?	| is.na(x)
!is.na() |	Is x not an NA? |	!is.na(x)



## [`arrange`](https://dplyr.tidyverse.org/reference/arrange.html) and `Pipe %>%`

* `arrange()` orders the rows of a data frame by the values of selected columns.

Unlike other `dplyr` verbs, `arrange()` largely ignores grouping; you need to explicitly mention grouping variables (`or use .by_group = TRUE) in order to group by them, and functions of variables are evaluated once per data frame, not once per group.

* [`pipes`](https://r4ds.had.co.nz/pipes.html) in R for Data Science.



## [`mutate`](https://dplyr.tidyverse.org/reference/mutate.html) 

* Create, modify, and delete columns

* Useful mutate functions

  - +, -, log(), etc., for their usual mathematical meanings

  - lead(), lag()

  - dense_rank(), min_rank(), percent_rank(), row_number(), cume_dist(), ntile()

  - cumsum(), cummean(), cummin(), cummax(), cumany(), cumall()

  - na_if(), coalesce()### `group_by()` and `summarise()`



## [`group_by`](https://dplyr.tidyverse.org/reference/group_by.html)



## [`summarise` or `summarize`](https://dplyr.tidyverse.org/reference/summarise.html)

#### Summary functions

So far our summarise() examples have relied on sum(), max(), and mean(). But you can use any function in summarise() so long as it meets one criteria: the function must take a vector of values as input and return a single value as output. Functions that do this are known as summary functions and they are common in the field of descriptive statistics. Some of the most useful summary functions include:

1. Measures of location - mean(x), median(x), quantile(x, 0.25), min(x), and max(x)
2. Measures of spread - sd(x), var(x), IQR(x), and mad(x)
3. Measures of position - first(x), nth(x, 2), and last(x)
4. Counts - n_distinct(x) and n(), which takes no arguments, and returns the size of the current group or data frame.
5. Counts and proportions of logical values - sum(!is.na(x)), which counts the number of TRUEs returned by a logical test; mean(y == 0), which returns the proportion of TRUEs returned by a logical test.


  - if_else(), recode(), case_when()



## Learn `dplyr` by Examples

### Data `iris`

```{r}
head(iris)
```




```{r}
summary(iris)
```



### `select` 1 - columns 1, 2, 5

```{r}
head(select(iris, c(1,2,5)))
```

You can select the first, the second and the fifth columns. If you want to use it, then assign a new name.

```{r}
head(iris)
```


```{r}
df_iris125 <- select(iris, c(1,2,5))
head(df_iris125)
```

### `select` 1 using pipe

In the previous example, we used `head(select(iris, c(1,2,5)))`, `head` comes first because we apply `head` to the result of `select(iris, c(1,2,5))`. In order to apply functions in a sequencial order, we can use pipe command. You can get the same result by the following.

```{r}
iris %>% select(c(1,2,5)) %>% head()
```

All `tidyverse` functions are designed so that the first argument, i.e., the entry, is the data. So using pipe, `iris` is assumed to be the first entry of the `select` function, and `select(iris, c(1,2,5))` is the first entry of the head function.

In the following, we use pipes.  

### `select` 2 - except Species

```{r}
select(iris, -Species) %>% head()
```



### `select` 3 - select and change column names at the same time

```{r}
select(iris, sl = Sepal.Length, sw = Sepal.Width, sp = Species) %>% head()
```

### `select` 4 - change the order of columns

```{r}
select(iris, c(5,3,4,1,2)) %>% head()
```


### `filter` - by names

```{r}
filter(iris, Species == "virginica") %>% head()
```




### `arrange`  - ascending and descending order

```{r}
arrange(iris, Sepal.Length, desc(Sepal.Width)) %>% head()
```



### `mutate` - rank

```{r}
iris %>% mutate(sl_rank = min_rank(Sepal.Length)) %>% 
  arrange(sl_rank) %>% head()
```

Insert a line break after the pipe command, not before. 

### `group_by` and `summarize`

```{r}
iris %>% 
  group_by(Species) %>% 
  summarize(sl = mean(Sepal.Length), sw = mean(Sepal.Width), 
  pl = mean(Petal.Length), pw = mean(Petal.Width))
```

* mean: `mean()` or `mean(x, na.rm = TRUE)` - arithmetic mean (average)
* median: `median()` or `median(x, na.rm = TRUE)` - mid value



For more examples see 

[dplr_iris](https://icu-hsuzuki.github.io/da4r2022_note/dplyr-iris.nb.html)


## References of `dplyr`

* Textbook: [R for Data Science, Part II Explore](https://r4ds.had.co.nz/wrangle-intro.html#wrangle-intro)


### RStudio Primers: See References in Moodle at the bottom

2. Work with Data -- [r4ds: Wrangle, I](https://r4ds.had.co.nz/wrangle-intro.html#wrangle-intro)
  - [Working with Tibbles](https://rstudio.cloud/learn/primers/2.1)
  - [Isolating Data with dplyr](https://rstudio.cloud/learn/primers/2.2)
  - [Deriving Information with dplyr](https://rstudio.cloud/learn/primers/2.3)



## Learn `dplyr` by Examples II - `gapminder`



### `ggplot2` [Overview](https://ggplot2.tidyverse.org)

`ggplot2` is a system for declaratively creating graphics, based on [The Grammar of Graphics](https://amzn.to/2ef1eWp). You provide the data, tell ggplot2 how to map variables to aesthetics, what graphical primitives to use, and it takes care of the details.

**Examples**
```
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```
```
ggplot(data = mpg) + 
  geom_boxplot(mapping = aes(x = class, y = hwy))
```

**Template**
```
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```



#### Gapminder and R Package `gapminder`

> Gapminder was founded by Ola Rosling, Anna Rosling Rönnlund, and Hans Rosling

-   Gapminder: <https://www.gapminder.org>

    -   Test on Top: You are probably wrong about - upgrade your worldview
    -   Bubble Chart: <https://www.gapminder.org/tools/#$chart-type=bubbles&url=v1>
    -   Dallar Street: <https://www.gapminder.org/tools/#$chart-type=bubbles&url=v1>
    -   Data: <https://www.gapminder.org/data/>

-   R Package gapminder by Jennifer Bryan

    -   Package site: <https://CRAN.R-project.org/package=gapminder>
    -   Site: <https://github.com/jennybc/gapminder>
    -   Documents: <https://www.rdocumentation.org/packages/gapminder/versions/0.3.0>

-   Package Help `?gapminder` or `gapminder` in the search window of Help

    -   The main data frame gapminder has 1704 rows and 6 variables:
        -   country: factor with 142 levels
        -   continent: factor with 5 levels
        -   year: ranges from 1952 to 2007 in increments of 5 years
        -   lifeExp: life expectancy at birth, in years
        -   pop: population
        -   gdpPercap: GDP per capita (US\$, inflation-adjusted)



```{r packages, message=FALSE}
library(tidyverse)
library(gapminder)
library(WDI)
```



#### R Package `gapminder` data

We will use a `tidyverse` function `slice` replacing `head`. Check `slice` in the search window under the Help tab on the bottom right pane.

```{r cache=TRUE}
df <- gapminder
df %>% slice(1:10)
```



```{r}
glimpse(df)
```



```{r}
summary(df)
```



#### Tidyverse::ggplot

##### First Try - with failures

You will encounter similar failures. We list three of them.

```{r}
ggplot(df, aes(x = year, y = lifeExp)) + geom_point()
```


There are lots of data in each year: 1952, 1957, 1962, 1967, 1972, 1977, 1982, 1987, 1992, 1997, …. Can you tell how many years are in the data? The following command shows different years in the data.

```{r}
unique(df$year)
```

You can guess it from the data summary above. Can you imagine how many countries are in the data? 142? Anyhow, too many points are on each year.

```{r}
ggplot(df, aes(x = year, y = lifeExp)) + geom_line()
```

Now, you can guess the reason why you had this output. This is often called a saw-tooth.

```{r}
ggplot(df, aes(x = year, y = lifeExp)) + geom_boxplot()
```

Can you see what the problem is? The `year` is a numerical variable in integer.

```{r}
typeof(pull(df, year)) # same as typeof(df$year)
```

The following looks better.

```{r}
ggplot(df, aes(y = lifeExp, group = year)) + geom_boxplot()
```



##### Box Plot

```{r}
ggplot(df, aes(x = as_factor(year), y = lifeExp)) + geom_boxplot()
```

We will study data visualization in Chapter \@ref(ggplot2). 

#### Applications of `dplyr`

Let us apply `dplyr` to manipulate data to visualize the data.

##### `filter`

By `filter` you can obtain the the data of one country.

`filter(country == "Afghanistan")`

Note that we need two equal symbols, and quotation marks must surround the country name.

```{r}
df %>% filter(country == "Afghanistan") %>%
  ggplot(aes(x = year, y = lifeExp)) + geom_line()
```

Looks good. From the data you observe, the life expectancy at birth in 1952 was below 30, and it was still below 44 in 2007.

Let us compare Afghanistan with Japan. When you choose more than one country, we use the following format: `country %in% c("Afghanistan", "Japan")`. 

```{r}
df %>% filter(country %in% c("Afghanistan", "Japan")) %>%
  ggplot(aes(x = year, y = lifeExp, color = country)) + geom_line()
```

What do you observe from this chart? 

The code `unique(df$country)` does the same as the one below. First, choose distinct elements in the column `country` by `distinct(country)` and get the column as a vector by `pull`.

```{r}
df %>% distinct(country) %>% pull()
```

As we have guessed, there are 142 countries in this data.

Let us choose BRICs countries in the data.


```{r}
df %>% filter(country %in% c("Brazil", "Russia", "India", "China")) %>%
  ggplot(aes(x = year, y = lifeExp, color = country)) + geom_line()
```

Russia data is missing. Can you find it in the list of countries? It can be a problem of `gapminder` data. Can you think of the reason why Russia is not in?



### Exercises

1.  Change `lifeExp` to `pop` and `gdpPercap` and do the same.
2.  Choose ASEAN countries and do the similar investigations.

-   Brunei, Cambodia, Indonesia, Laos, Malaysia, Myanmar, Philippines, Singapore.

- How many of these countries are on the list?

3.  Choose several countries by yourself and do the similar investigations.



### `group_by` and `summarize`

Let us use the variable `continent` and summarize the data. Can you tell how many continents are listed in the data? Yes, there are five. Can you tell how many countries are in each continent on the data?

```{r}
df_lifeExp <- df %>% group_by(continent, year) %>% 
  summarize(mean_lifeExp = mean(lifeExp), median_lifeExp = median(lifeExp), max_lifeExp = max(lifeExp), min_lifeExp = min(lifeExp), .groups = "keep")
```

Don't get scared. We will learn little by little. 


```{r}
df_lifeExp %>% slice(1:10)
```

You can use `fill` and `color` for the box plot. Try and check the difference.

```{r}
df %>% filter(year %in% c(1952, 1987, 2007)) %>%
  ggplot(aes(x=as_factor(year), y = lifeExp, fill = continent)) +
  geom_boxplot()
```

The following are examples of line graphs. Please see the differences.

```{r}
df_lifeExp %>% ggplot(aes(x = year, y = mean_lifeExp, color = continent)) +
  geom_line()
```



```{r}
df_lifeExp %>% ggplot(aes(x = year, y = mean_lifeExp, color = continent, linetype = continent)) +
  geom_line()
```



```{r}
df_lifeExp %>% ggplot() +
  geom_line(aes(x = year, y = mean_lifeExp, color = continent)) + 
  geom_line(aes(x = year, y = median_lifeExp, linetype = continent))
```


## The Week Two Assignment (in Moodle)

**R Markdown and `dplyr`**

* Create an R Notebook of a Data Analysis containing the following and submit the rendered HTML file (eg. `a2_123456.nb.html`)
  1. create an R Notebook using the R Notebook Template in Moodle,  save as `a2_123456.Rmd`, 
  2. write your name and ID and the contents, 
  3. run each code block, 
  4. preview to create `a2_123456.nb.html`,
  5. submit  `a2_123456.nb.html` to Moodle.

1. Pick data from the built-in datasets besides `cars`. (`library(help = "datasets")` or go to the site [The R Datasets Package](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/00Index.html))

    - Information of the data: Name, Description, Usage, Format, Source, References (Hint: ?cars)
    - Use `head()`, `str()`, ..., and create at least one chart using `ggplot2` - Code Chunk.
      + Don't forget to add `library(tidyverse)` in the first code chunk.
    - An observation of the chart - in your own words.



2. Load `gapminder` by `library(gapminder)`.

    - Choose `pop` or `gdpPercap`, or both, one country in the data, a group of countries in the data.
    - Create charts using ggplot2 with geom_line and the variables and countries chosen in 1. (See examples of the charts for `lifeExp`.)
    - Study the data as you like.
    - Observations and difficulties encountered.

**Due:** 2023-01-09 23:59:00. Submit your R Notebook file in Moodle (The Second Assignment). Due on Monday!



### Original Data? WDI?

```{r}
gapminder %>% slice(1:10)
```



#### WDI

* SP.DYN.LE00.IN: Life expectancy at birth, total (years)
* NY.GDP.PCAP.KD: GDP per capita (constant 2015 US$)
* SP.POP.TOTL: Population, total

```{r eval=FALSE}
df_wdi <- WDI(
  country = "all", 
  indicator = c(lifeExp = "SP.DYN.LE00.IN", pop = "SP.POP.TOTL", gdpPercap = "NY.GDP.PCAP.KD")
)
```
```{r echo=FALSE, eval=FALSE}
df_wdi <- WDI(
  country = "all", 
  indicator = c(lifeExp = "SP.DYN.LE00.IN", pop = "SP.POP.TOTL", gdpPercap = "NY.GDP.PCAP.KD")
)
write_csv(df_wdi, "./data/wdi.csv")
```
```{r echo=FALSE, eval=TRUE}
df_wdi <- read_csv("./data/wdi.csv")
```



```{r}
df_wdi %>% slice(1:10)
```



```{r eval=FALSE}
df_wdi_extra <- WDI(
  country = "all", 
  indicator = c(lifeExp = "SP.DYN.LE00.IN", pop = "SP.POP.TOTL", gdpPercap = "NY.GDP.PCAP.KD"), 
  extra = TRUE
)
```
```{r eval=FALSE, echo=FALSE}
df_wdi_extra <- WDI(
  country = "all", 
  indicator = c(lifeExp = "SP.DYN.LE00.IN", pop = "SP.POP.TOTL", gdpPercap = "NY.GDP.PCAP.KD"), 
  extra = TRUE
)
write_csv(df_wdi_extra, "./data/wdi_extra.csv")
```
```{r eval=TRUE, echo=FALSE}
df_wdi_extra <- read_csv("./data/wdi_extra.csv")
```



```{r}
df_wdi_extra
```

Can you see the differences? List them out. We will study the World Development Indicators in Chapter \@ref(wdi). 
