
# Responses to Assignment Two

1. You are supposed to submit an R Notebook File with a file name a2_YourID.nb.html.
  - Some submitted an HTML file, such as a2_YourID.html. You need to create an R Notebook. Use the template in Moodle. It creates a file with *.nb.html at the end automatically.
  - Some did not run each code chunk. You should run each code or select 'Run all' under 'Run' button. If some code chunk has a problem or an error, run each code chunk or use Run all chunk above or Run all chunk below, so the result appear in your R Notebook file.
  
2. You are supposed to write observations. 
  - Writing codes seem to be challenging, however, we are learning 'data analysis' not 'programming'. Do not forget to write explanations of the data, questions and observations.
  
3. Cheat Sheets, Posit Primers, and the textbook 'R for Data Science' are the first set of references you should look at together wih my lecture materials.

# Set up

```{r}
library(tidyverse)
library(gapminder)
```

The following `(df <- gapminder)` is a short-hand of 

```
df <- gapminder
df
```

```{r}
(df <- gapminder)
```

# General Comments

## Varibles

We should know first about the variables. At least you must know if each of the variables is a categorical variable or a numerical variable.

For example, in the `gapminder` data, `country`, `continent` are categorical variables, and `year`, `lifeExp`, `pop`, `gdpPercap` are numerical variables. It is possible to treat `year` as a categorical variable. 

## Example: `datasets::CO2`

### The first step

You can obtain basic information of the data by the following or typing CO2 in the search box under Help tab. You can see the same at: https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/00Index.html

```{r eval=FALSE}
help(CO2) # or ? CO2
```

* **Description**: The CO2 data frame has 84 rows and 5 columns of data from an experiment on the cold tolerance of the grass species Echinochloa crus-galli.
* **Usage**: CO2
* **Format**
  - An object of class c("nfnGroupedData", "nfGroupedData", "groupedData", "data.frame") containing the following columns:

  - Plant: an ordered factor with levels Qn1 < Qn2 < Qn3 < ... < Mc1 giving a unique identifier for each plant.

  - Type: a factor with levels Quebec Mississippi giving the origin of the plant

  - Treatment: a factor with levels nonchilled chilled

  - conc: a numeric vector of ambient carbon dioxide concentrations (mL/L).

  - uptake: a numeric vector of carbon dioxide uptake rates (\mu\mbox{mol}/m^2μmol/m 2
  sec).
  
* **Details**: The CO_2 uptake of six plants from Quebec and six plants from Mississippi was measured at several levels of ambient CO_2 concentration. Half the plants of each type were chilled overnight before the experiment was conducted.
  - This dataset was originally part of package nlme, and that has methods (including for [, as.data.frame, plot and print) for its grouped-data classes.

* **Source**: Potvin, C., Lechowicz, M. J. and Tardif, S. (1990) “The statistical analysis of ecophysiological response curves obtained from experiments involving repeated measures”, Ecology, 71, 1389–1400.
  - Pinheiro, J. C. and Bates, D. M. (2000) Mixed-effects Models in S and S-PLUS, Springer.


```{r}
df_co2 <- as_tibble(datasets::CO2) # what happens if simply `df_co2 <- datasets::CO2`
df_co2
```

You can use `head(CO2)` if you set `df_co2 <-CO2` or `df_co2 <- datasets::CO2`.

```{r}
glimpse(df_co2)
```

"factor" is a categorical data, and "double" is a numerical data.

```{r}
class(df_co2$Plant)
class(df_co2$Type)
class(df_co2$Treatment)
class(df_co2$conc)
class(df_co2$uptake)
```
```{r}
summary(df_co2)
```

### Try as many visualizations as possible

Then you can choose appropriate ones later in your research.

#### Histogram


```{r}
df_co2 %>% ggplot(aes(x = conc)) + geom_histogram()
```

```{r}
df_co2 %>% ggplot(aes(x = uptake)) + geom_histogram()
```

#### Box Plots

```{r}
df_co2 %>% ggplot(aes(x = factor(conc), y = uptake)) + geom_boxplot()
```

```{r}
df_co2 %>% ggplot(aes(x = factor(conc), y = uptake, fill = Type)) + geom_boxplot()
```
```{r}
df_co2 %>% ggplot(aes(x = factor(conc), y = uptake, fill = Treatment)) + geom_boxplot()
```


```{r}
df_co2 %>% ggplot(aes(x = Plant, y = uptake, fill = Treatment)) + geom_boxplot()
```


```{r}
df_co2 %>% ggplot(aes(x = Plant, y = uptake, fill = Type)) + geom_boxplot()
```


```{r}
df_co2 %>% ggplot(aes(x = Plant, y = uptake, fill = Type, color = Treatment)) + geom_boxplot()
```

What can you see?
Write your observations.

#### Scatter Plots

```{r}
df_co2 %>% ggplot(aes(x = conc, y = uptake, color = Treatment)) + geom_point()
```


```{r}
df_co2 %>% ggplot(aes(x = Plant, y = Type, color = Treatment, size = conc)) + geom_point()
```


```{r}
df_co2 %>% ggplot(aes(x = Plant, y = Type, size = conc, shape = Treatment)) + geom_point() + facet_wrap(vars(Treatment))
```


```{r}
ggplot(data = df_co2) + 
  geom_point(aes(x = conc, y = uptake))
```

The following prints a vector.

```{r}
df_co2 %>% distinct(conc) %>% pull()
```

The following code generates a data frame.

```{r}
df_co2 %>% distinct(conc)
```

```{r}
ggplot(data = CO2) + 
  geom_line(aes(x = conc, y = uptake))
```

The code above did not work, and the line graph is not appropriate in this case. There are so many `update` values at the same `conc`.

### Example. `datasets::Seatbelts`

Search the data information.

**Road Casualties in Great Britain 1969–84**

* Seatbelts is a multiple time series, with columns
* DriversKilled: car drivers killed.
* drivers: same as UKDriverDeaths.
* front: front-seat passengers killed or seriously injured.
* rear: rear-seat passengers killed or seriously injured.
* kms: distance driven.
* PetrolPrice: petrol price.
* VanKilled: number of van (‘light goods vehicle’) drivers.
* law: 0/1: was the law in effect that month?

**References**
Harvey, A. C. and Durbin, J. (1986). The effects of seat belt legislation on British road casualties: A case study in structural time series modelling. Journal of the Royal Statistical Society series A, 149, 187–227. doi:10.2307/2981553.

_The paper is available as you log-in to ICU Library > E-Databases > JSTOR_

Can you see the difference of the following two codes?

```{r}
head(Seatbelts)
```


```{r}
df_sb <- as_tibble(datasets::Seatbelts)
df_sb
```

```{r}
summary(df_sb)
```

Which visualization do you apply?

```{r}
df_sb %>% ggplot(aes(x = factor(law), y = DriversKilled)) + geom_boxplot()
```

What do you observe above?

```{r}
df_sb %>% ggplot(aes(x = PetrolPrice, y = DriversKilled)) + geom_point() +
  geom_smooth(formula = y~x, method = "lm", se = FALSE)
```

What can you see above?

```{r}
df_sb %>% ggplot(aes(x = kms, y = DriversKilled)) + geom_point() +
  geom_smooth(formula = y~x, method = "lm", se = FALSE)
```

What can you see above?

We will learn how to use `pivot_longer` and `pivot_wider` in EDA4.

```{r}
df_sb %>% 
  pivot_longer(cols = 2:4, names_to = "seat", values_to = "value")
```

```{r}
df_sb %>% 
  pivot_longer(cols = 2:4, names_to = "seat", values_to = "value") %>%
  ggplot() + geom_boxplot(aes(x = seat, y = value, fill = seat))
```

What can you observe?



```{r}
df %>% filter(continent == "Asia") %>%
  ggplot(aes(x = year, y = gdpPercap, color = country)) + geom_line()
```

Appropriate graph?

## Gapminder

```{r}
ggplot(df, aes(x = continent)) + geom_bar()
```

```{r}
ggplot(df, aes(x = continent, fill = continent)) + geom_bar(width = 0.75)
```

```{r}
ggplot(df, aes(x = continent, y = pop)) + geom_boxplot()
```

```{r}
ggplot(df, aes(x = continent, y = log10(pop))) + geom_boxplot()
```

Alternately, you can use `coord_trans(x = "identity", y = "log10")` in stead of `y = log10(pop)`.
Can you see the difference?

```{r}
ggplot(df, aes(x = continent, y = pop)) + geom_boxplot() + 
  coord_trans(x = "identity", y = "log10")
```

```{r}
df %>% filter(year %in% c(1957, 1982, 2007)) %>%
  ggplot() + geom_boxplot(aes(x = continent, y = pop, fill = factor(year))) +
    coord_trans(x = "identity", y = "log10")
```


```{r}
df %>% filter(year %in% c(1957, 1982, 2007)) %>%
  ggplot() + geom_boxplot(aes(x = continent, y = gdpPercap, fill = factor(year))) +
  coord_trans(x = "identity", y = "log10")
```


```{r}
ggplot(df, aes(gdpPercap, lifeExp)) + geom_point(aes(color=continent))
```


```{r}
df %>% ggplot(aes(gdpPercap, lifeExp, color = continent)) + geom_point() +
  coord_trans(x = "log10", y = "identity")
```

```{r}
df %>% filter(year %in% c(2007)) %>%
  ggplot() + geom_point(aes(x = pop, y = gdpPercap, col = continent)) +
  coord_trans(x = "log10", y = "log10")
```
```{r}
df %>% filter(year %in% c(1957)) %>%
  ggplot() + geom_point(aes(x = gdpPercap, y = lifeExp, col = continent, size = pop)) +
  coord_trans(x = "log10", y = "identity")
```

```{r}
df %>% filter(year %in% c(2007)) %>%
  ggplot() + geom_point(aes(x = gdpPercap, y = lifeExp, col = continent, size = pop)) +
  coord_trans(x = "log10", y = "identity")
```

```{r}
df %>% filter(year %in% c(1957, 2007)) %>%
  ggplot() + geom_point(aes(x = gdpPercap, y = lifeExp, col = continent, size = pop)) +
  coord_trans(x = "log10", y = "identity") +
  facet_wrap(vars(year))
```



```{r}
df_lifeExp <- df %>% group_by(continent, year) %>% 
  summarize(mean_lifeExp = mean(lifeExp), median_lifeExp = median(lifeExp), max_lifeExp = max(lifeExp), min_lifeExp = min(lifeExp))
```

The code above gives a message, but it works.

```{r}
df_lifeExp %>% ggplot(aes(x = year, y = median_lifeExp, color = continent)) +
  geom_line()
```

If you do not want to have a message, the following is an option. Otherwise, grouping is kept and you can get the original data back by `ungroup()`.

```{r}
df_pop <- df %>% group_by(continent, year) %>% 
  summarize(mean_pop = mean(pop), median_pop = median(pop), max_pop = max(pop), min_pop = min(pop), .groups = "drop")
```

```{r}
df_pop %>% ggplot(aes(x = year, y = mean_pop, color = continent)) +
  geom_line()
```
The following two codes create the same chart.

```{r}
df_pop %>% ggplot(aes(x = year, y = mean_pop, color = continent, linetype = continent)) +
  geom_line()
```

```{r}
df_pop %>% ggplot() +
  geom_line(aes(x = year, y = mean_pop, color = continent)) + 
  geom_line(aes(x = year, y = median_pop, linetype = continent))
```



# Responses to Questions

## Q1. Two categorical variables and one numerical variables

Eg. [Smoking, Alcohol and (O)esophageal Cancer](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/esoph.html)

```{r}
(df_esoph <- as_tibble(esoph))
```

`df_esoph` has three categorical variables and one numerical variable `ncases` to investigate.

**Comments**: I wanted to include three variables in the first exercise to be able to compare tobacco consumption, number of cases of cancer, and age in the same graph but I was not able to do it.

**Solutions**: There are various ways you can choose from.

Scatter plot with size by `geom_point()`.

```{r}
ggplot(df_esoph) + geom_point(aes(agegp, tobgp, size=ncases))
```

Heatmap with `geom_tile()`

```{r}
ggplot(df_esoph) + geom_tile(mapping = aes(x = agegp, y = tobgp, fill = ncases))
```

`geom_boxplot()`

```{r}
ggplot(df_esoph, aes(x= tobgp, y=ncases, fill=agegp))+geom_boxplot()
```

```{r}
ggplot(df_esoph, aes(x= agegp, y=ncases, fill=tobgp))+geom_boxplot()
```

`geom_col()`

```{r}
ggplot(df_esoph, aes(x= agegp, y=ncases, fill=tobgp))+geom_col()
```

Default position is "stack".

```{r}
ggplot(df_esoph, aes(x= agegp, y=ncases, fill=tobgp))+geom_col(position = "dodge")
```


## Q2. Combine two charts

```{r}
df %>%
  select(country, continent, year, gdpPercap) %>%
  filter(continent %in% c("Asia", "Europe")) %>%
  group_by(continent, year) %>%
  summarise(mean_GDPperCapita = mean(gdpPercap)) %>%
  ggplot(aes(x=year)) +
  geom_line(aes(y=mean_GDPperCapita, color=continent)) +
  ggtitle("GDP oer capita by continents, 1950's to today") 
```


```{r}
df %>% 
  select(country, year, gdpPercap) %>% 
  filter(country %in% c("Israel", "Japan", "Norway", "China", "Ireland")) %>% 
  ggplot(aes(x=year)) + 
  geom_line(aes(y=gdpPercap, color=country))
```

**Question.** I have not managed to add on the same graph of the continents the data for the individual countries, as I would have liked:

**Solution.** Construct two data sets and combine them into one.

`ggplot2` starts with one data. 

```{r}
df_2c <- df %>%
  select(continent, year, gdpPercap) %>%
  filter(continent %in% c("Asia", "Europe")) %>%
  group_by(continent, year) %>%
  summarise(gdpPercap = mean(gdpPercap), .groups = 'drop') %>%
  select(country = continent, year, gdpPercap)

df_5c <- df %>%
  select(country, year, gdpPercap) %>%
  filter(country %in% c("Israel", "Japan", "Norway", "China", "Ireland"))

df_2c %>% bind_rows(df_5c) %>%
  ggplot(aes(x = year, y = gdpPercap, color = country)) + geom_line()
```

Use `mutate`.

```{r}
df %>%
  group_by(continent, year) %>%
  mutate(mean_by_continent = mean(gdpPercap)) %>%
  ungroup() %>%
  filter(country %in% c("Israel", "Japan", "Norway", "China", "Ireland")) %>%
  ggplot(aes(x = year)) +
    geom_line(aes(y = gdpPercap, color=country)) +
    geom_line(aes(y = mean_by_continent, linetype=continent)) +
    labs(title = "GDP oer capita of five countries", subtitle = "Mean of GDP per capita of their continent") 
```

When you want to change the linetype manually, use `scale_linetype_manual()`.

```{r}
df %>%
  group_by(continent, year) %>%
  mutate(mean_by_continent = mean(gdpPercap)) %>%
  ungroup() %>%
  filter(country %in% c("Israel", "Japan", "Norway", "China", "Ireland")) %>%
  ggplot(aes(x = year)) +
    geom_line(aes(y = gdpPercap, color=country)) +
    geom_line(aes(y = mean_by_continent, linetype=continent)) + 
    scale_linetype_manual(values = c("Asia" = "twodash", "Europe" = "longdash")) + 
    labs(title = "GDP oer capita of five countries", subtitle = "Mean of GDP per capita of their continent") 
```

# Appendix: Change colors, shapes, linetypes, etc. manually

Example: Default

```{r}
df %>%
  filter(country %in% c("Germany", "Japan", "United States")) %>%
  ggplot() +
    geom_line(aes(x = year, y = gdpPercap, color=country, linetype=continent))
```

* `scale_color_manual`: https://ggplot2-book.org/scale-colour.html
  - eg1: scale_colour_manual(values = c("red", "blue", "green"))
  - eg2: scale_colour_manual(values = c("China" = "red", "Japan" = "blue", "Norway" = "green"))
  - eg3: scale_colour_manual(values = scales::hue_pal()(3)) # default
  - eg4: scale_colour_manual(values = scales::hue_pal(direction = -1)(3)) # reverse order
* `scale_fill_manual`: similar to  `scale_color_manual`
* `scale_linetype_manual`: https://ggplot2-book.org/scale-other.html?q=linetype#scale-linetype
* `scale_shape_manual: https://ggplot2-book.org/scale-other.html?q=scale_shape_manual#scale-shape
* `scale_size: https://ggplot2-book.org/scale-other.html?q=size#scale-size


```{r}
df %>%
  filter(country %in% c("Germany", "Japan", "United States")) %>%
  ggplot(aes(x = year, y = gdpPercap)) +
    geom_line(aes(color=country, linetype=continent)) +
    geom_point(aes(shape = country)) +
    scale_colour_manual(values = scales::hue_pal(direction = -1)(3)) +
    scale_linetype_manual(values = c("Europe" = "dotted", "Asia" = "dotdash", "Americas" = "longdash")) +
    scale_shape_manual(values = c("Germany" = 7, "Japan" = 9, "United States" = 12))
```