# Importing Public Data, WDI {#wdi}

## Reviews and Previews

```{r message=FALSE}
library(tidyverse)
library(gapminder)
library(maps)
library(WDI)
library(readxl)
library(ggrepel)
```

* We have used `tidyverse` and `gapminder` already.
* If you have not installed `WDI`, install it.
* We will not use `ggrepel` but if you want to use it, install it.
* `maps` and `readxl` are bundled in `tidyverse` but need to be attached by `library`.

---

### Gapminder Package Data

```{r}
df <- gapminder
df
```

---

### `gdpPercap` of ASEAN countries

```{r eval=FALSE}
asean <- c("Brunei", "Cambodia", "Laos", "Myanmar", 
           "Philippines", "Indonesia", "Malaysia", "Singapore")
df %>% filter(country %in% asean) %>%
  ggplot(aes(x = year, y = gdpPercap, col = country)) + geom_line()
```

---

```{r echo=FALSE}
asean <- c("Cambodia", "Myanmar", 
           "Philippines", "Indonesia", "Malaysia", "Singapore")
df %>% filter(country %in% asean) %>%
  ggplot(aes(x = year, y = gdpPercap, col = country)) + geom_line()
```


---

```{r echo=TRUE}
df %>% filter(country %in% asean) %>%
  ggplot(aes(x = gdpPercap, y = lifeExp, col = country)) + geom_point()
```

---

```{r echo=TRUE}
df %>% filter(country %in% asean) %>%
  ggplot(aes(x = gdpPercap, y = lifeExp, col = country)) + 
  geom_point() + coord_trans(x = "log10", y = "identity")
```

$\log_{10}{100}$ = `r log10(100)`, $\log_{10}{1000}$ = `r log10(1000)`, $\log_{10}{10000}$ = `r log10(10000)`

<!-- $10^{2.5}$ = `r sqrt(10)^{5}`, $10^{3}$ = `r 10^{3}`, $10^{3.5}$ = `r sqrt(10)^7`,  -->

<!-- $10^{4}$ = `r 10^4`, $10^{4.5}$ = `r sqrt(10)^9`. -->

---

```{r gapminder-combined, eval=FALSE}
library(ggrepel)
df2007 <- df %>% filter(country %in% asean, year == 2007)
df %>% filter(country %in% asean) %>%
  ggplot(aes(x = gdpPercap, y = lifeExp, col = country))+ 
  geom_line() + geom_label_repel(data = df2007, aes(label = country)) + geom_point()  +
  coord_trans(x = "log10", y = "identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1), legend.position = "none") +
  labs(title = "Life Expectancy vs GDP Per Capita of ASEAN Countries",
       subtitle = "Data: gapminder package", x = "GDP per Capita", y = "Life Expectancy")
```

---


```{r echo=FALSE}
library(ggrepel)
df2007 <- df %>% filter(country %in% asean, year == 2007)
df %>% filter(country %in% asean) %>%
  ggplot(aes(x = gdpPercap, y = lifeExp, col = country))+ 
  geom_line() + geom_label_repel(data = df2007, aes(label = country)) + geom_point()  +
  coord_trans(x = "log10", y = "identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1), legend.position = "none") +
  labs(title = "Life Expectancy vs GDP Per Capita of ASEAN Countries",
       subtitle = "Data: gapminder package", x = "GDP per Capita", y = "Life Expectancy")
```

---


```{r echo=FALSE}
library(tidyverse)
library(maps)
world_map <- map_data("world")
df %>%
  ggplot(aes(map_id = country)) + 
  geom_map(aes(fill = gdpPercap), map = world_map) + expand_limits(x = world_map$long, y = world_map$lat) +
  labs(title = "Gapminder Package Data", subtitle="World Map of GDP per Capita Data")
```

---

### World Bank: World Development Indicators (WDI)

* SP.DYN.LE00.IN: Life expectancy at birth, total (years)
* NY.GDP.PCAP.KD: GDP per capita (constant 2015 US$)
* SP.POP.TOTL: Population, total

```{r cash=TRUE, eval=FALSE}
df_wdi <- WDI(
  country = "all", 
  indicator = c(lifeExp = "SP.DYN.LE00.IN", pop = "SP.POP.TOTL", gdpPercap = "NY.GDP.PCAP.KD")
)
```

```{r echo=FALSE, message=FALSE, eval=FALSE}
write_csv(df_wdi, "data/df_wdi.csv")
```

```{r echo=FALSE, message=FALSE}
df_wdi <- read_csv("data/df_wdi.csv")
```


---

```{r}
df_wdi
```

---

```{r cash=TRUE, eval=FALSE}
df_wdi_extra <- WDI(
  country = "all", 
  indicator = c(lifeExp = "SP.DYN.LE00.IN", pop = "SP.POP.TOTL", gdpPercap = "NY.GDP.PCAP.KD"), 
  extra = TRUE
)
```

```{r echo=FALSE, message=FALSE, eval=FALSE}
write_csv(df_wdi_extra, "data/df_wdi_extra.csv")
```

```{r echo=FALSE, message=FALSE}
df_wdi_extra <- read_csv("data/df_wdi_extra.csv")
```


---

```{r}
df_wdi_extra
```


---

## Exploratory Data Analysis

### What is EDA (Posit Primers: [Visualise Data](https://posit.cloud/learn/primers/3.1))

1. EDA is an iterative cycle that helps you understand what your data says. When you do EDA, you:

2. Generate questions about your data

3. Search for answers by visualising, transforming, and/or modeling your data

Use what you learn to refine your questions and/or generate new questions

EDA is an important part of any data analysis. You can use EDA to make discoveries about the world; or you can use EDA to ensure the quality of your data, asking questions about whether the data meets your standards or not.

---

## Open and Public Data, World Bank

### [Open Government Data Toolkit](http://opendatatoolkit.worldbank.org): [Open Data Defined](http://opendatatoolkit.worldbank.org/en/essentials.html)

The term **Open Data** has a very precise meaning. Data or content is open if anyone is free to use, re-use or redistribute it, subject at most to measures that preserve provenance and openness.

1. The data must be _legally open_, which means they must be placed in the public domain or under liberal terms of use with minimal restrictions.
2. The data must be _technically open_, which means they must be published in electronic formats that are machine readable and non-proprietary, so that anyone can access and use the data using common, freely available software tools. Data must also be publicly available and accessible on a public server, without password or firewall restrictions. To make Open Data easier to find, most organizations create and manage Open Data catalogs.

---

## World Bank: WDI - World Development Indicaters

* World Bank: https://www.worldbank.org
* [Who we are](https://www.worldbank.org/en/who-we-are):
  - To end extreme poverty: By reducing the share of the global population that lives in extreme poverty to 3 percent by 2030.
  - To promote shared prosperity: By increasing the incomes of the poorest 40 percent of people in every country. 
* World Bank Open Data: https://data.worldbank.org
  - Data Bank, World Development Indicators, etc.
* [World Development Indicators (WDI)](https://datatopics.worldbank.org/world-development-indicators/) : the World Bank’s premier compilation of cross-country comparable data on development; 1400 time series indicators
  - Themes: Poverty and Inequality, People, Environment, Economy, States and Markets, Global Links
  - Open Data & DataBank: Explore data, Query database
  - Bulk Download: Excel, CSV
  - API Documentation
  
---

## R Package [WDI](https://CRAN.R-project.org/package=WDI)

* [WDI](https://CRAN.R-project.org/package=WDI): World Development Indicators and Other World Bank Data
* Search and download data from over 40 databases hosted by the World Bank, including the World Development Indicators ('WDI'), International Debt Statistics, Doing Business, Human Capital Index, and Sub-national Poverty indicators.
* Version: 2.7.4
* Materials:	[README](https://cran.r-project.org/web/packages/WDI/readme/README.html)   - _usage_
  - [NEWS](https://cran.r-project.org/web/packages/WDI/news/news.html) - _version history_
* Published: 2021-04-06
* README: https://cran.r-project.org/web/packages/WDI/readme/README.html
* Reference manual:	[WDI.pdf](https://cran.r-project.org/web/packages/WDI/WDI.pdf)

---

## Function WDI

* **Usage**

```
WDI(country = "all",
    indicator = "NY.GDP.PCAP.KD",
    start = 1960,
    end = 2020,
    extra = FALSE,
    cache = NULL)
```

* **Arguments** See Help!
  - country: Vector of countries (ISO-2 character codes, e.g. "BR", "US", "CA", or "all") 
  - indicator: If you supply a named vector, the indicators will be automatically renamed: `c('women_private_sector' = 'BI.PWK.PRVS.FE.ZS')`

---

## Function WDIsearch

```{r}
library(WDI)
```
```{r}
WDIsearch(string = "NY.GDP.PCAP.KD", 
          field = "indicator", cache = NULL)
```

---

```{r}
WDIsearch(string = "population", 
          field = "name", short=FALSE, cache = NULL)
```

---

```
WDIsearch(string = "NY.GDP.PCAP.KD", 
  field = "indicator", short = FALSE, cache = NULL)
```
```
WDIsearch(string = "gdp", 
  field = "name", short = TRUE, cache = NULL) 
```

---

## Bulk Downloads at WDI site

WDIbulk downloads the zip file of Bulk Downloads in [WDI site](https://datatopics.worldbank.org/world-development-indicators/) , it is a list containing 6 data frames: Data, Country, Series, Country-Series, Series-Time, FootNote.

---

## WDIcache

Download an updated list of available WDI indicators from the World Bank website. Returns a list for use in the WDIsearch function.

```{r widcache, cash=TRUE, eval=FALSE}
wdi_cache <- WDIcache()
```

Downloading all series information from the World Bank website can take time. The WDI package ships with a local data object with information on all the series available on 2012-06-18. You can update this database by retrieving a new list using `WDIcache`, and then feeding the resulting object to `WDIsearch` via the cache argument.

```{r echo=FALSE, message=FALSE, eval=FALSE}
write_rds(wdi_cache, "data/wdi_cache.RData", refhook = NULL)
```

```{r echo=FALSE, message=FALSE, eval=FALSE}
wdi_cache <- read_rds("data/wdi_cache.RData", refhook = NULL)
```



---

```{r eval=FALSE}
wdi_cache
```


---

## WDI_data

List of 2 data frames

The first character matrix includes a full list of WDI series. This list is updated semi-regularly. Users can refresh the list manually using the 'WDIcache()' function and search in the updated list using the 'cache' argument.




```{r}
WDI_data$country  %>% filter(country == "Japan")
```

---

```{r}
WDIsearch(string = "gdp", 
  field = "name", short = FALSE, cache = wdi_cache) 
```

---

## World Development Indicators - Summary

Find indicators:

1. `WDIsearch(string = "gdp", field = "name", short = FALSE, cache = NULL)`
  - `WDIsearch(string = "gdp", field = "name", short = FALSE, cache = wdi_cache)`
  - `WDIsearch(string = "NY.GDP.PCAP.KD", field = "indicator", short = FALSE, cache = NULL)`
2. [WDI](https://datatopics.worldbank.org/world-development-indicators/): Data Themes
3. Browse by Indicators: https://data.worldbank.org/indicator
   - Featured Indicators or All Indicators
   - Obtain the indicator from the detail or the URL

---

### Example: CO2 emissions (metric tons per capita)

* ID: EN.ATM.CO2E.PC
* URL: https://data.worldbank.org/indicator/EN.ATM.CO2E.PC

```{r}
WDIsearch(string = "EN.ATM.CO2E.PC", field = "indicator", 
          short = FALSE, cache = wdi_cache) 
```

```{r}
WDIsearch(string = "EN.ATM.CO2E.PC", field = "indicator", 
          short = FALSE, cache = wdi_cache) %>% pull(description) 
```

* Source: Climate Watch. 2020. GHG Emissions. Washington, DC: World Resources Institute. Available at: climatewatchdata.org/ghg-emissions. See SP.POP.TOTL for the denominator's source.




---

```{r cash=TRUE, eval=FALSE}
co2pcap <- WDI(country = "all", indicator = "EN.ATM.CO2E.PC", start = 1960, end = NULL, extra = TRUE, cache = wdi_cache)
```

```{r echo=FALSE, message=FALSE, eval=FALSE}
write_csv(co2pcap, "data/co2pcap.csv")
```

```{r echo=FALSE, message=FALSE}
co2pcap <- read_csv("data/co2pcap.csv")
```

```{r}
co2pcap
```


---

```{r}
write_csv(co2pcap, "data/co2pcap.csv")
```


---

```{r}
co2pcap %>% filter(country %in% c("World", "Japan", "United States", "China")) %>%
  ggplot(aes(x = year, y = EN.ATM.CO2E.PC, color = country)) + 
  geom_line()
```

---

```{r}
co2pcap %>% filter(!is.na(EN.ATM.CO2E.PC)) %>% pull(year) %>% summary()
```




---

```{r}
co2pcap %>% 
  filter(country %in% c("World", "Japan", "United States", "China"), year %in% 1990:2019) %>%
  ggplot(aes(x = year, y = EN.ATM.CO2E.PC, color = country)) + 
  geom_line()
```

---

```{r}
co2pcap %>% 
  filter(income != "Aggregates", year == 2019) %>%
  ggplot(aes(x = income, y = EN.ATM.CO2E.PC, fill = income)) + 
  geom_boxplot()
```

---

```{r}
co2pcap %>% 
  filter(income != "Aggregates", year == 2019, !is.na(EN.ATM.CO2E.PC)) %>%
  ggplot(aes(x = income, y = EN.ATM.CO2E.PC, fill = income)) + 
  geom_boxplot()
```

* What is `boxplot`: https://vimeo.com/222358034

---

```{r}
co2pcap %>% 
  filter(income != "Aggregates", year == 2019, !is.na(EN.ATM.CO2E.PC)) %>%
  group_by(income) %>%
  summarize(min = min(EN.ATM.CO2E.PC), med = median(EN.ATM.CO2E.PC), max = max(EN.ATM.CO2E.PC), IQR = IQR(EN.ATM.CO2E.PC), n = n())
```

---
```{r}
co2pcap %>% 
  filter(income != "Aggregates", year == 2019, !is.na(EN.ATM.CO2E.PC)) %>%
  filter(!income %in% c("High income", "Low income", "Lower middle income", "Upper middle income"))
```

```{r}
co2pcap %>% 
  filter(income != "Aggregates", year == 2019) %>%
  filter(income == "Not classified")
```

---

```{r echo=TRUE, eval=FALSE}
co2pcap %>% 
  filter(income != "Aggregates", year == 2019) %>%
  ggplot(aes(map_id = country)) + 
  geom_map(aes(fill = income), map = world_map) + expand_limits(x = world_map$long, y = world_map$lat) +
  labs(title = "Income Levels in 2019")
```
---

```{r echo=FALSE}
co2pcap %>% 
  filter(income != "Aggregates", year == 2019) %>%
  ggplot(aes(map_id = country)) + 
  geom_map(aes(fill = income), map = world_map) + expand_limits(x = world_map$long, y = world_map$lat) +
  labs(title = "Income Levels in 2019")
```

---

```{r}
co2pcap %>% distinct(country)
```

---

```{r}
world_map %>% distinct(region)
```

---

```{r}
world_map0 <- world_map %>% 
  mutate(region = case_when(region == "Macedonia" ~ "North Macedonia",
                            region == "Ivory Coast"  ~ "Cote d'Ivoire",
                            region == "Democratic Republic of the Congo"  ~ "Congo, Dem. Rep.",
                            region == "Republic of Congo" ~  "Congo, Rep.",
                            region == "UK" ~  "United Kingdom",
                            region == "USA" ~  "United States",
                            region == "Laos" ~  "Lao PDR",
                            region == "Slovakia" ~  "Slovak Republic",
                            region == "Saint Lucia" ~  "St. Lucia",
                            region == "Kyrgyzstan"  ~  "Kyrgyz Republic",
                            region == "Micronesia" ~ "Micronesia, Fed. Sts.",
                            region == "Swaziland"  ~ "Eswatini", 
                            region == "Virgin Islands"  ~ "Virgin Islands (U.S.)", 
                            region == "Russia" ~ "Russian Federation", 
                            region == "Egypt" ~ "Egypt, Arab Rep.",
                            region == "South Korea" ~ "Korea, Rep.",
                            region == "North Korea" ~ "Korea, Dem. People's Rep.",
                            region == "Iran" ~ "Iran, Islamic Rep.",
                            region == "Brunei" ~ "Brunei Darussalam",
                            region == "Venezuela" ~ "Venezuela, RB",
                            region == "Yemen" ~ "Yemen, Rep.",
                            region == "Bahamas" ~ "Bahamas, The",
                            region == "Syria" ~ "Syrian Arab Republic",
                            region == "Turkey" ~ "Turkiye",
                            region == "Cape Verde" ~ "Cabo Verde",
                            region == "Gambia" ~ "Gambia, The",
                            region == "Czech Republic" ~ "Czechia",
                            TRUE ~ region))
```

---

```{r}
write_csv(world_map0, "data/world_map0.csv")
```

```{r eval=FALSE}
map0_url <- "https://icu-hsuzuki.github.io/da4r2022_note/data/world_map0.csv"
world_map0 <- read_csv(map0_url)
```


---

```{r}
co2pcap %>% filter(income != "Aggregates", year == 2019) %>% 
  anti_join(world_map0, by = c("country"="region"))
```

---

```{r}
world_map0 %>% anti_join(co2pcap, by = c("region"="country")) %>% distinct(region) %>% arrange(region)
```

---



```{r}
world_map0 %>% left_join(iso3166, by = c("region" = "ISOname")) %>%
  filter(is.na(a2)) %>% distinct(region)
```


---


```{r echo=FALSE}
co2pcap %>% 
  filter(income != "Aggregates", year == 2019) %>%
  ggplot(aes(map_id = country)) + 
  geom_map(aes(fill = income), map = world_map0) + expand_limits(x = world_map$long, y = world_map$lat) +
  labs(title = "Income Levels in 2019")
```


