
# Assignment Three {#a3resp}

### Assignment Three: The Week Four Assignment {-}

**WDI and `ggplot2`**

* Create an R Notebook of a Data Analysis containing the following and submit the rendered HTML file (eg. `a3_123456.nb.html`  by replacing 123456 with your ID)
  1. create an R Notebook using the R Notebook Template in Moodle,  save as `a3_123456.Rmd`, 
  2. write your name and ID and the contents, 
  3. run each code block, 
  4. preview to create `a3_123456.nb.html`,
  5. submit  `a3_123456.nb.html` to Moodle.

1. Choose at least one indicator of WDI

    - Information of the data: Name, Indicator, Description, Source, etc.
    - Download the data with `WDI`
    - Explain why you chose the indicator
    - List questions you want to study

2. Explore the data using visualization using `ggplot2`

    - Use a histogram (geom_histogram), boxplot (geom_boxplot), a scatter plot (geom_point), a line plot (geom_line)
    - For at least one chart, add title, and labels of axis, and add an explanation of it

3. Observations and difficulties encountered.

**Due:** 2023-01-16 23:59:00. Submit your R Notebook file in Moodle (The Third Assignment). Due on Monday!

## Set up

Follow the workflow explained in EDA4 on January 18.

### EDA by R Studio: Step 1

In RStudio,

1.1. Project

  * Create a new project: File > New Project; or  
  * Open a project: File > Open Project, Open Project in New Session, Open Recent Project  
    - It is easier to find an existing project from: File > Recent Project 
  * _Check there is a file `project_name.Rproj` in your project folder (directory)_

 
1.2. data folder (directory) `data`

  * Create a data folder: Press New Folder at the right bottom pane; or 
  * Confirm the data folder previously created: Press Files at the right bottom pane
  * _If you follow 1, the data folder exists in your project folder_

 
1.3. Move (or copy) data for the project to the data folder

  * If you downloaded the data, it is in your Download folder. Move it to `data`.
  * _Check in your RStudio that your data is in `data`: Press Files at the right bottom pane and click `data`, the data folder._


### EDA by R Studio: Step 2

2.1. Project Notebook: Memo

  - Create an R Notebook: File > New File > R Notebook
  
    + You can use R Notebook template in Moodle by moving the template (template.Rmd or template.nb.Rmd) file in your project folder or copy and paste the text file into your new R Notebook.
    + If you use template.nb.Rmd (R Notebook File), choose Open in Editor.
    
  - Add descriptive title. 
  
2.2. Setup Code Chunk 

  - Create a code chunk and add packages to use in the project and RUN the code.
  
      + library(tidyverse)
      + library(WDI)
      + or any other packages
      
```{r 9-1}
library(tidyverse)
library(WDI)
```

_When people read a paper, they want to know the minimal set of packages required to install. So it is better to load the packages by library() you actually need in the paper._

2.3. Choose `Source` or `Visual` editor mode, and start editing Project Notebook

  - Set up Headings such as: About, Data, Analysis and Visualizations, Conclusions
  - Under About or Data, paste url of the sites and/or the data

2.4. Edit a new file by saving as for a report

  - File > Save As...

## General Comments

### Varibles

We should know first about the variables. At least you must know if each of the variables is a categorical variable or a numerical variable.

### Visualization

  * What type of variation occurs within my variables?
  * What type of covariation occurs between my variables?

## World Development Indicators (WDI)

### Setup

It is not a must, but it is good to run the following code chunk and use `wdi_cache` for `WDIsearch()` to update the mata data. See `WDIcache` help. 

```{r 9-2, eval=FALSE}
wdi_cache <- WDIcache()
```
```{r 9-3, echo=FALSE, eval=FALSE}
wdi_cache <- WDIcache()
write_rds(wdi_cache, "./data/wdi_cache.RData")
```
```{r 9-4, echo=FALSE, eval=TRUE}
wdi_cache <- read_rds("./data/wdi_cache.RData")
```

### Indicators Used in Assignment Three

* NY.GDP.PCAP.KD: GDP per capita (constant 2015 US$)
* NY.GDP.MKTP.CD: GDP (current US$)
* NY.GDP.PCAP.KD.ZG: GDP per capita growth (annual%)
* SP.POP.TOTL: Total Population
* IP.JRN.ARTC.SC: Scientific and technical journal articles
* FP.CPI.TOTL.ZG: Inflation, consumer prices (% annual growth) 
* DT.ODA.ODAT.PC.ZS: Net ODA received per capita (current US$)
* BX.KLT.DINV.CD.WD: Foreign direct investment, net inflows (BoP, current US$)
* MS.MIL.XPND.GD.ZS: Military expenditure (% of GDP)
* MS.MIL.TOTL.P1: Total Armed Forces Personnel
* SI.POV.MDIM: Multidimensional poverty headcount ratio (% of total population)
* SI.POV.NAHC: Poverty headcount ratio at national poverty lines (% of population) 
* SI.POV.GINI: Gini index
* NY.GNS.ICTR.ZS: Gross savings (% of GDP)
* SL.UEM.TOTL.ZS: Unemployment, total (% of total labor force) (modeled ILO estimate)
* SP.URB.TOTL.IN.ZS (Urban population (% of total population))
* AG.LND.AGRI.ZS: Agricultural land (% of land area)
* SE.ADT.LITR.FE.ZS: Literacy rate, adult female (% of females ages 15 and above))
* SE.ADT.LITR.MA.ZS: Literacy rate, adult male (% of males ages 15 and above)) 
* SE.ADT.LITR.ZS: Literacy rate, adult total
* SE.ADT.1524.LT.FE.ZS: Literacy rate, youth female (% of females ages 15-24)
* SE.XPD.TOTL.GD.ZS: Government expenditure on education, total (% of GDP)
* SG.GEN.PARL.ZS: Proportion of seats held by women in national parliaments (%)

### WDIsearch

We obtain the information of the data, i.e., meta data of WDI: Name, Indicator, Description, Source, etc. by `WDIsearch`.

* Usage

```
WDIsearch(string = "gdp", field = "name", short = TRUE, cache = NULL)
```

* Arguments

  * `string`: Character string. Search for this string using grep with ignore.case=TRUE.

  * `field`: Character string. Search this field. Admissible fields: 'indicator', 'name', 'description', 'sourceDatabase', 'sourceOrganization'

  * `short`: TRUE: Returns only the indicator's code and name. FALSE: Returns the indicator's code, name, description, and source.

  * `cache`: Data list generated by the WDIcache function. If omitted, WDIsearch will search a local list of series.

If you follow the order of the arguments, you can omit argument names: `string`, `field`, `short`, `cache`.

* The simplest format 1.

```{r 9-5}
WDIsearch("NY.GNS.ICTR.ZS", "indicator")
```
* The simplest format 2.

```{r 9-6}
WDIsearch("Literacy rate", "name")
```

* Detailed with short = FALSE

```{r 9-7}
WDIsearch("IP.JRN.ARTC.SC", "indicator", FALSE)
```
* You can use `wdi_cache` downloaded above to use the updated meta data.

```{r 0-8}
WDIsearch("Government expenditure on education", "name", FALSE, wdi_cache)
```


### Importing Data Using WDI

* The shortest form. The following is same as:
`WDI(country = "all", indicator = "NY.GDP.PCAP.KD", start = 1960, end = NULL, extra = FALSE, cache = NULL, latest = NULL, language = "en)`

```{r 9-9, eval=FALSE}
WDI(indicator = "NY.GDP.PCAP.KD")
```
```{r 9-10, echo=FALSE, eval=FALSE}
df_gdp <- WDI(indicator = "NY.GDP.PCAP.KD")
write_csv(df_gdp, "./data/gdp.csv")
```
```{r 9-11, echo=FALSE, eval=TRUE}
df_gdp <- read_csv("./data/gdp.csv")
df_gdp
```

* In order to use the downloaded data, we need to assign a name to it. It can be `df` if you download and use only one data, but it is better to assign the data to be a more descriptive name.

```{r 9-12, eval=FALSE}
df_gdppcap <- WDI(indicator = "NY.GDP.PCAP.KD")
df_gdppcap
```
```{r 9-13, echo=FALSE, eval=FALSE}
df_gdppcap <- WDI(indicator = "NY.GDP.PCAP.KD")
write_csv(df_gdp, "./data/gdppcap.csv")
```
```{r 9-14, echo=FALSE, eval=TRUE}
df_gdppcap <- read_csv("./data/gdppcap.csv")
df_gdppcap
```

* If you want to use extra information such as `region`, `income`, `lending`, use `extra = TRUE`.

```{r 9-15, eval=FALSE}
df_gdppcap_extra <- WDI(indicator = "NY.GDP.PCAP.KD", extra = TRUE)
df_gdppcap_extra
```
```{r 9-16, echo=FALSE, eval=FALSE}
df_gdppcap_extra <- WDI(indicator = "NY.GDP.PCAP.KD", extra = TRUE)
write_csv(df_gdppcap_extra, "./data/gdppcap_extra.csv")
```
```{r 9-17, echo=FALSE, eval=TRUE}
df_gdppcap_extra <- read_csv("./data/gdppcap_extra.csv")
df_gdppcap_extra
```

### Writing and reading a csv file

Since the data is generally huge, it is better to use a data saved in your computer.
Before saving it check whether you have `data` folder (or directory) in you project folder. Use the `Files` tab of the right bottom pane. You should be able to see the project icon with Rproj at the end, your R Notebook file you are editing, and the `data` folder. Since `.csv` is automatically added, `write_csv(gdppcap_extra, "./data/gdppcap_extra")` does the same as below.

```{r 9-18, eval = FALSE}
write_csv(df_gdppcap_extra, "./data/gdppcap_extra.csv")
```

To read the data, run the next code chunk. You do not have to run the code `df_gdppcap_extra <- WDI(indicator = "NY.GDP.PCAP.KD", extra = TRUE)` again.

```{r 9-19, eval = FALSE}
gdppcap <- read_csv("./data/gdppcap_extra.csv")
```

"./data/gdpcap_extra.csv" is the way to express the name of the data `gdpcap_extra.csv` in csv format in the data folder. 

### `country` argument

If you want to import data for several countries, you can use `iso2c` codes of the countries.

```{r 9-20}
ASEAN <- c("BN", "ID", "KH", "LA", "MM", "MY", "PH", "SG")
```

```{r 9-21, eval=FALSE}
df_gdppcap_asean <- WDI(ASEAN, "NY.GDP.PCAP.KD")
```
```{r 9-22, echo=FALSE, eval=FALSE}
df_gdppcap_asean <- WDI(ASEAN, "NY.GDP.PCAP.KD")
write_csv(df_gdppcap_asean, "./data/gdppcap_asean.csv")
```
```{r 9-23, echo=FALSE, eval=TRUE}
df_gdppcap_asean <- read_csv("./data/gdppcap_asean.csv")
df_gdppcap_asean
```

```{r 9-24}
df_gdppcap_asean %>% distinct(country) %>% pull()
```


```{r 9-25}
wdi_cache$country %>% filter(iso2c %in% ASEAN) %>% distinct(country) %>% pull()
```


You can also use `wdi_cache$country`.

```{r 9-26}
wdi_cache$country %>% slice(1:10)
```

```{r 9-27}
wdi_cache$country %>% filter(iso2c %in% ASEAN) %>% pull(country)
```


You can find iso2c codes from the downloaded data, or using wdi_cache$country

```{r 9-28}
wdi_cache$country %>% 
  filter(country %in% c("Brunei Darussalam", "Indonesia", "Cambodia", "Lao PDR", "Myanmar", "Malaysia", "Philippines", "Singapore")) %>%
  pull(iso2c)
```

You can also find countries in some category.

```{r 9-29}
wdi_cache$country %>% 
  filter(region == "South Asia") %>%
  pull(country)
```


```{r 9-30}
wdi_cache$country %>% 
  filter(income == "Lower middle income") %>%
  pull(iso2c)
```

The following shows a list of indicators

```{r 9-31}
wdi_cache$series %>% slice(1:10)
```

The following is same as `WDIsearch("gdp per cap", "name")`. See WDIsearch help.

```{r 9-32, eval=FALSE}
wdi_cache$series %>% filter(grepl("gdp per cap", name, ignore.case = TRUE))
```

## Responses to Questions

### Q1. How do we use histograms.

* [Histograms and frequency polygons](https://ggplot2.tidyverse.org/reference/geom_histogram.html)
* [Posit Cloud: Histograms](https://posit.cloud/learn/primers/3.3)


```{r 9-33}
wdi_cache$series %>% filter(indicator == "SG.GEN.PARL.ZS") %>% pull(name)
```
```{r 9-34, eval=FALSE}
df_women_in_parl <- WDI(indicator = c(women_in_parl = "SG.GEN.PARL.ZS")) %>%
  drop_na(women_in_parl)
```

```{r 9-35, echo=FALSE, eval=FALSE}
write_csv(df_women_in_parl, "./data/women_in.parl.csv")
```
```{r 9-36, echo=FALSE, eval=TRUE}
df_women_in_parl <-read_csv("./data/women_in.parl.csv")
```

```{r 9-37, }
df_women_in_parl %>% filter(year == 2020) %>%
  ggplot(aes(women_in_parl)) + geom_histogram()
```

```{r 9-38}
df_women_in_parl %>% filter(year == 2020) %>%
  ggplot(aes(women_in_parl)) + geom_histogram(binwidth = 5)
```

```{r 9-39}
wdi_cache$series %>% filter(indicator == "IP.JRN.ARTC.SC") %>% pull(name)
```

```{r 9-40, eval=FALSE}
df_stja <- WDI(indicator = c(stja = "IP.JRN.ARTC.SC"), extra = TRUE) %>%
  drop_na(stja)
```
```{r 9-41, echo=FALSE, eval=FALSE}
write_csv(df_stja, "./data/stja.csv")
```
```{r 9-42, echo=FALSE, eval=TRUE}
df_stja <-read_csv("./data/stja.csv")
```


```{r 9-43}
df_stja$year %>% summary()
```

Default of the number of bins is 30. Change and find an appropriate one.


```{r 9-44}
df_stja %>% filter(income != "Aggregates") %>% 
  ggplot(aes(stja)) + geom_histogram(bins = 20)
```
```{r 9-45}
df_stja %>% filter(income != "Aggregates", stja >0) %>% 
  ggplot(aes(stja)) + geom_histogram(bins = 20) + scale_x_log10()
```
```{r 9-46}
df_stja %>% filter(income != "Aggregates", income != "Not classified", stja >0) %>% 
  ggplot(aes(stja, fill = income)) + geom_histogram(alpha = 0.7, bins = 20, color = "black") + scale_x_log10()
```

```{r 9-47}
df_stja %>% filter(income != "Aggregates", income != "Not classified", stja >0) %>% 
  ggplot(aes(stja, fill = income)) + geom_density(alpha = 0.3) + scale_x_log10()
```

```{r 9-48}
wdi_cache$series %>% filter(indicator == "MS.MIL.XPND.GD.ZS") %>% pull(name)
```


```{r 9-49, eval=FALSE}
df_milxpnd <- WDI(indicator = c(milxpnd = "MS.MIL.XPND.GD.ZS"), extra = TRUE) %>%
  drop_na(milxpnd)
```
```{r 9-50, echo=FALSE, eval=FALSE}
write_csv(df_milxpnd, "./data/milxpnd.csv")
```
```{r 9-51, echo=FALSE, eval=TRUE}
df_milxpnd <-read_csv("./data/milxpnd.csv")
```

Default of the number of bins is 30. Here I chose binwidth = 0.5 (%).

```{r 9-52}
df_milxpnd %>% filter(income != "Aggregates", year == 2021) %>% 
  ggplot(aes(milxpnd, fill = income)) + geom_histogram(color = "black", binwidth = 0.5) + 
  labs(title = "Military expenditure (% of GDP)")
```

```{r 9-53}
df_milxpnd %>% filter(income != "Aggregates", income != "Not classified", year == 2021) %>% 
  ggplot(aes(milxpnd, fill = income)) + geom_density(binwidth = 0.5, alpha = 0.3) + 
  labs(title = "Military expenditure (% of GDP)")
```

### Q2. Effect of unemployed rate under pandemic depending on income groups

```{r 9-54}
wdi_cache$series %>% filter(indicator == "SL.UEM.TOTL.ZS") %>% pull(name)
```

```{r 9-55, eval=FALSE}
df_ur <- WDI(indicator = c(ur = "SL.UEM.TOTL.ZS"), 
             extra = TRUE, cache = wdi_cache) %>% drop_na(ur)
```

```{r 9-56, echo=FALSE, eval=FALSE}
write_csv(df_ur, "./data/ur.csv")
```
```{r 9-57, echo=FALSE, eval=TRUE}
df_ur <-read_csv("./data/ur.csv")
```

```{r 9-58}
df_ur %>% filter(income == "Aggregates") %>% filter(grepl('income', country)) %>% 
  filter(year >= 2018) %>% ggplot() + geom_line(aes(x = year, y = ur, color = country))
```

### Q3. Add values at each data point.

```{r 9-59}
wdi_cache$series %>% filter(indicator == "FP.CPI.TOTL.ZG") %>% pull(name)
```

```{r 9-60, eval=FALSE}
df_infl <- WDI(country = c("VN","CN","JP"), 
               indicator = c(inf = "FP.CPI.TOTL.ZG"), extra = TRUE) %>% drop_na(inf)
```
```{r 9-61, echo=FALSE, eval=FALSE}
write_csv(df_infl, "./data/infl.csv")
```
```{r 9-62, echo=FALSE, eval=TRUE}
df_infl <-read_csv("./data/infl.csv")
```
```{r 9-63}
df_infl_s <- df_infl %>% filter(year %in% c(2000, 2005, 2010, 2015, 2020), 
                   country %in% c("Japan", "Vietnam", "China"))
df_infl_s
```
```{r 9-64}
df_infl_s %>%
  ggplot(aes(x = year, y = inf, color= country)) + 
  geom_line() + geom_point() + 
  geom_text(aes(label = round(inf,2)), nudge_y = 0.8) + 
labs(title = "Inflation of Japan, Vietnam and China from 2000 to 2020", 
     x = "", y = "Inflation, consumer prices (annual %)")
```

## Regression Lines; Topic of EDA5

```{r 9-65}
wdi_cache$series %>% filter(indicator %in% c("SI.POV.NAHC","SI.POV.MDIM")) %>% pull(name)
```

```{r 9-66, cache=TRUE, eval=FALSE}
df_wdi_poverty <- WDI(
  indicator = c(poverty = "SI.POV.NAHC", multipoverty = "SI.POV.MDIM", gdppercap = "NY.GDP.PCAP.KD"), start = 1990,
  extra = TRUE) %>% drop_na(poverty, multipoverty, gdppercap)
```

```{r 9-67, echo=FALSE, eval=FALSE}
write_csv(df_wdi_poverty, "./data/wdi_poverty.csv")
```
```{r 9-68, echo=FALSE, eval=TRUE}
df_wdi_poverty <-read_csv("./data/wdi_poverty.csv")
```

```{r 9-69}
df_wdi_poverty %>%  
  group_by(country, year) %>%
  mutate(mean_gdp = mean(gdppercap)) %>%
  mutate(mean_poverty= mean(poverty)) %>%
  ungroup() %>% filter(income != "Aggregates") %>% 
  ggplot(aes(x = mean_gdp)) + geom_point(aes(y = mean_poverty, color = income)) +
  scale_x_log10() +  geom_smooth(aes(y = mean_poverty), formula = y~x, linetype="longdash", color = "black", method = "lm", se = FALSE) + 
  labs(x = "GDP per capita", y = "poverty rate (% of population)", title = "Poverty rates and GDP per capita", subtitle="world countries, 1990-2021 average, by income level")
```

```{r 9-70}
df_wdi_poverty %>%  
  group_by(country, year) %>%
  mutate(mean_gdp = mean(gdppercap)) %>%
  mutate(mean_multipoverty= mean(multipoverty)) %>%
  ungroup() %>%
    filter(region !="Aggregates") %>% ggplot(aes(x = mean_gdp)) + geom_point(aes(y = mean_multipoverty, color = region)) +  
  scale_x_log10() +  
  geom_smooth(aes(y = mean_multipoverty), formula = y~x, linetype="longdash", color = "black", method = "lm", se = FALSE) + labs(x = "GDP per capita", y = "Multidimentinal poverty rate (% of population)", title = "Multidimentional Poverty rates and GDP per capita", subtitle="world countries, 1990-2021 average, by region")
```


## Appendix: A secondary axis

* [Specify a secondary axis](https://ggplot2.tidyverse.org/reference/sec_axis.html)
  - This function is used in conjunction with a position scale to create a secondary axis, positioned opposite of the primary axis. All secondary axes must be based on a one-to-one transformation of the primary axes.

`scale_y_continuous(sec.axis = sec_axis(~ . scaling_function))`

### Example

Suppose you have two indicators, 

* NY.GDP.MKTP.KD: GDP (constant 2015 US$)
* NY.GDP.PCAP.KD: GDP per capita (constant 2015 US$)

```{r 9-71}
WDIsearch(string = "NY.GDP.MKTP.KD", field = "indicator", short = FALSE, cache = wdi_cache)
```

```{r 9-72}
WDIsearch(string = "NY.GDP.PCAP.KD", field = "indicator", short = FALSE, cache = wdi_cache)
```

List the name of countries of ASEAN and BRICs using `wdi_cache$country`.

```{r 9-73}
asean <- c("Brunei Darussalam", "Cambodia", "Lao PDR", "Myanmar", 
           "Philippines", "Indonesia", "Malaysia", "Singapore")
brics <- c("Brazil", "Russian Federation", "India", "China")
```

Find the `iso2c` of the countries using `wdi_cache$country`.

```{r 9-74}
wdi_cache$country %>% 
  filter(country %in% 
           c("Brunei Darussalam", "Cambodia", "Lao PDR", "Myanmar", 
           "Philippines", "Indonesia", "Malaysia", "Singapore",
           "Brazil", "Russian Federation", "India", "China")) %>% 
  pull(iso2c)
```

Separate the iso3c's of the countries with commas and read data using `WDI`.

```{r 9-75, eval=FALSE}
wdi_gdp <- WDI(
  country = c("BR", "BN", "CN", "ID", "IN", "KH", "LA", "MM", "MY", "PH", "RU", "SG"),
  indicator = c(gdp = "NY.GDP.MKTP.KD", gdpPercap = "NY.GDP.PCAP.KD"),
  start = 1960, extra = TRUE, cache = wdi_cache)
```

```{r 9-76, echo=FALSE, eval=FALSE}
write_csv(wdi_gdp, "./data/wdi_gdp.csv")
```
```{r 9-77, echo=FALSE, eval=TRUE}
wdi_gdp <-read_csv("./data/wdi_gdp.csv")
```

```{r 9-78}
wdi_gdp %>% filter(country %in% asean) %>% drop_na(gdp, gdpPercap) %>% summary() 
```

Using the summary, decide the scaling of two variables, `gdpPercap` and `gdp`.

```{r 9-79}
wdi_gdp %>% drop_na(gdp, gdpPercap) %>%
  filter(country %in% asean) %>%
  ggplot() + 
  geom_line(aes(x = year, y = gdpPercap, linetype = country)) + 
  geom_line(aes(x = year, y = gdp/(10^7), col = country)) + 
  coord_trans(x ="identity", y="log10") +
  scale_y_continuous(sec.axis = sec_axis(~ . *(10^7), name = "gdp/(10^7)"))
```

#### BRICs

```{r 9-80}
wdi_gdp %>% filter(country %in% brics) %>% drop_na(gdp, gdpPercap) %>% summary() 
```

Using the summary, decide the scaling of two variables, `gdpPercap` and `gdp`.

```{r 9-81}
wdi_gdp %>% drop_na(gdp, gdpPercap) %>%
  filter(country %in% brics) %>%
  ggplot() + 
  geom_line(aes(x = year, y = gdpPercap, linetype = country)) + 
  geom_line(aes(x = year, y = gdp/(10^9), col = country)) + 
  coord_trans(x ="identity", y="log10") +
  scale_y_continuous(sec.axis = sec_axis(~ . *(10^7), name = "gdp/(10^9)"))
```


