# Our World in Data {#owid}


## The package `owidR`

> We study the R package `owidR` for importing data from Our World in Data.

## References

The package official site contains other links. When you quote the package, use the link to the official site.

* Package Official Site: https://CRAN.R-project.org/package=owidR
  - ReadMe: https://cran.r-project.org/web/packages/owidR/readme/README.html
  - Manual: https://cran.r-project.org/web/packages/owidR/owidR.pdf
  - Vignette: [Create and Analyse a Dataset](https://cran.r-project.org/web/packages/owidR/vignettes/example-analysis.html)
* [owidR: Import Data from Our World in Data](https://rdrr.io/cran/owidR/)
  - Man page and source codes
  
In general, README gives a short introduction to the package, a Manual, the comprehensive descriptions of each function, and a Vignette, a practical introduction containing examples and applications.

## Introduction

>This package acts as an interface to [Our World in Data](https://ourworldindata.org) datasets, allowing for an easy way to search through data used in over 3,000 charts and load them into the R environment.

### Setup

#### Installation

Run the following for the first time

```{r eval=FALSE}
install.packages("owidR")
```

#### Load the package

```{r}
library(owidR)
library(tidyverse) 
```

The package automatically load a part of `tidyverse`, e.g., `dplyr`, `ggplot`, .... Since it works well with the scheme`tidyverse`, it is better to load `tidyverse` with it.

The creator of this package also suggests loading packages `plm` for panels of data, and `texreg` for displaying models, but let us start without them until we actually use them. For panel data, see, for example, the site](https://www.aptech.com/blog/introduction-to-the-fundamentals-of-panel-data/).

### Core functions

#### List of core functions

In this package, `chart` is close to data, and `chart id` is a data indicator.

* owid: Get a dataset used in an OWID chart
* owid_covid: Get the Our World in Data covid-19 dataset
* owid_search: Search the data sources used in OWID charts
* owid_source: A function to get source information from an OWID dataset and display it in the R console.
* pal_owid: Colour palettes based on the colours used by Our World in Data
* view_chart: A function that opens the original OWID chart in your browser
* world_map_data: Function that returns a simple feature collection of class sf. Map data is from naturalearthdata.com. Designed to be used internally.

#### Usage

##### `owid_search`

Search the data sources used in OWID charts

```
owid_search(term)
```

* Example

Since the output is long, I cut it off to the first six rows using `head()`.

```{r}
owid_search("emissions") %>% head()
```

A matrix is returned. If the list is long, it is easier to see the pairs of the titles and chart_ids by adding `as_tibble()`.

```{r}
owid_search("emissions") %>% as_tibble()
```

If the list is not long, you do not need to add `as_tibble()`. However, note that you need to keep in mind that the title and the chart_id consists of a pair, and you need to use the chart_id to download the data using `owid`.

```{r}
owid_search("human rights")
```


##### `owid`

Get a dataset used in an OWID chart

```
owid(chart_id = NULL, rename = NULL, tidy.date = TRUE, ...)
```

`chard_id`: The chart_id as returned by owid_search, which is combined with '-'. Don't mix up with the chart titles.

`rename`: Rename the value column. Currently only works if their is just one value col- umn.

* Example

```{r cash=TRUE}
emissions <- owid("per-capita-ghg-emissions")
```

```{r}
emissions
```

```{r cash=TRUE}
rights <- owid("human-rights-scores")
```

```{r}
rights
```

**Note.**

1. You can use `rename` to change column names. For example, 

```{r}
owid("per-capita-ghg-emissions", rename = "ghgPcap")
```

2. Since until after importing the data, you never know the original column name, and how many columns are for indicators. It is natural to change column names using `dpyr::rename`. In the next example, I used `Total including LUCF`. However, 'Total including LUCF' and "Total including LUCF" work as well.

```{r}
emissions %>% rename(ghgPcap = `Total including LUCF`)
```

3. If there are more than one variables to rename, use vector notation as follows. Here `top_n(1)` is same as `slice(1)`, and gives the first row only.

```{r}
owid("electoral-democracy") %>% top_n(1)
```

```{r}
owid("electoral-democracy", rename = c("electoral_democracy", "vdem_high", "vdem_low"))
```


4. You can use `dplyr::rename`, and keep the record of renaming column names.

```{r}
(democracy <- owid("electoral-democracy"))
```


```{r}
democracy <- democracy %>% 
  rename(`electoral_democracy` = `Electoral democracy`,  
         `vdem_high` = `electdem_vdem_high_owid`, 
         `electdem_vdem_low_owid` = `electdem_vdem_low_owid`)
democracy
```


##### `owid_source`

A function to get source information from an OWID dataset and display it in the R console.

```
owid_source(data)
```

* Example

```{r}
owid_source(emissions)
```


```{r}
owid_source(rights)
```



##### `view_chart`

A function that opens the original OWID chart in your browser

```
view_chart(x)
```

* `x` Either a tibble returned by `owid()`, or a `chart_id`.

* Example

The first one uses the chart, i.e., the tibble returned by `owid()`, and the second, `chart_id`. You can also embed in your R Markdown file by copying Embed `iframe` clink from `Share` botton at the bottom right corner.

```{r eval=FALSE}
firearm_suicide <- owid("suicide-rate-by-firearm")
    view_chart(firearm_suicide)
```

<iframe src="https://ourworldindata.org/grapher/suicide-rate-by-firearm" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>

```{r eval=FALSE}
view_chart("electoral-democracy")
```

<iframe src="https://ourworldindata.org/grapher/electoral-democracy?country=ARG~AUS~BWA~CHN" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>

```{r eval=FALSE}
view_chart("share-of-individuals-using-the-internet")
```

<iframe src="https://ourworldindata.org/grapher/share-of-individuals-using-the-internet" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>


##### `pal_owid`

Colour palettes based on the colours used by Our World in Data

##### `owid_covid`

owid_covid: Get the Our World in Data covid-19 dataset

```
owid_covid()
```

See the detail at the [GitHub site](https://github.com/owid/covid-19-data/tree/master/public/data#data-on-covid-19-coronavirus-by-our-world-in-data).

* Example

```{r cash=TRUE}
covid <- owid_covid()
```

```{r}
covid %>% filter(location == "Japan")
```

## Examples

The following is based on the presentation and the first two R Notebook files created by Professor Kaizoji.

### Human Rights- modified from  README

Lets use the core functions to get data on how human rights have changed over time. First by searching for charts on human rights.

```{r}
owid_search("human rights") %>% as_tibble()
```

Let’s use the human rights protection dataset.

```{r}
rights <- owid("human-rights-protection")
rights
```

ggplot2 makes it easy to visualise our data.


```{r}
rights %>% 
  filter(entity %in% c("United Kingdom", "France", "United States", "Japan")) %>%  
  ggplot(aes(year, `Human rights protection`, colour = entity)) +
  geom_line()
```


### Internet - modified from vignette


```{r}
owid_search("internet") %>% as_tibble()
```

Get a dataset used in an OWID chart.

```{r}
owid("share-of-individuals-using-the-internet") %>% top_n(1)
```

```{r}
internet <- owid("share-of-individuals-using-the-internet", rename = "internet_use")
internet
```

Get source information on an OWID dataset

```{r}
owid_source(internet)
```

A function that opens the original OWID chart in your browser.

```{r eval=FALSE}
view_chart(internet)
```

<iframe src="https://ourworldindata.org/grapher/share-of-individuals-using-the-internet" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>

Plot an owid dataset.  The first is the simplest, and the second uses oied theme.

```{r}
internet %>% filter(entity == "World") %>%
  ggplot(aes(year, internet_use))+ geom_line() 
```

```{r}
internet %>% filter(entity == "World") %>%
  ggplot(aes(year, internet_use))+ geom_line() +
  labs(title = "Share of the World Population \nusing the Internet",
       y = "Individuals using the Internet \n(% of population)") +
  scale_y_continuous(limits = c(0, 100))
```

```{r}
internet %>% 
  filter(entity %in% c("United Kingdom", "Spain", "Russia", "Egypt", "Nigeria")) %>% 
  ggplot(aes(year, internet_use, color =  entity)) + geom_line() +
  labs(title = "Share of Population with Using the Internet",
       y = "Individuals using the Internet \n(% of population)",
       color = "country") +
  scale_y_continuous(limits = c(0, 100), labels = scales::label_number(suffix = "%"))
```

Creating a choropleth map.

```{r}
owid_map(internet, year = 2017) +
  labs(title = "Share of Population Using the Internet, 2017")
```

### Democracy - modified from vignette

```{r}
owid_search("democrac") %>% as_tibble()
```

```{r}
owid("electoral-democracy") %>% top_n(1)
```

```{r}
democracy <- owid("electoral-democracy", rename = c("electoral_democracy", "vdem_high", "vdem_low"))
democracy
```

```{r}
owid_source(democracy)
```

```{r}
owid_map(democracy, year = 2015, palette = "YlGn") +
  labs(title = "Electoral Democracy")
```

```{r}
democracy %>% 
  filter(entity %in% c("United Kingdom", "Spain", "Russia", "Egypt", "Nigeria")) %>%
  ggplot(aes(year, electoral_democracy, color = entity)) + geom_line() +
  labs(title = "Electoral Democracy", y = "", color = "country") +
  scale_y_continuous(limits = c(0, 1), labels = scales::label_number(suffix = "%"))
```


```{r}
gdp <- owid("gdp-per-capita-worldbank", rename = "gdp")

gov_exp <- owid("total-gov-expenditure-gdp-wdi", rename = "gov_exp")

age_dep <- owid("age-dependency-ratio-of-working-age-population", rename = "age_dep")

unemployment <- owid("unemployment-rate", rename = "unemp")
```

Mutating joins

left_join(): includes all rows in x.

**References**

* See https://dplyr.tidyverse.org/reference/mutate-joins.html.
* Posit Primers - Tidy your data: https://posit.cloud/learn/primers/4
  - Join Data Sets: https://posit.cloud/learn/primers/4.3

```{r}
data <- internet %>% 
  left_join(democracy) %>% 
  left_join(gdp) %>% 
  left_join(gov_exp) %>% 
  left_join(age_dep) %>% 
  left_join(unemployment)
```

Drawing scatter plot

```{r message=FALSE, warning=FALSE}
data %>% 
  filter(year == 2015) %>% 
  ggplot(aes(internet_use, electoral_democracy)) +
  geom_point(colour = "#57677D", na.rm = TRUE) +
  geom_smooth(method = "lm", colour = "#DC5E78", na.rm = TRUE) +
  labs(title = "Relationship Between Internet Use \nand electoral_democracy", x = "Internet Use", y = "electoral_democracy")
```


```{r message=FALSE, warning=FALSE}
data %>% 
  filter(year == 2015) %>% 
  ggplot(aes(gdp, internet_use)) +
  geom_point(colour = "blue") +
  geom_smooth(method = "gam", colour = "red", level = 0.0) +
  labs(title = "Relationship Between Internet Use and GDP", x = "GDP", y = "Internet Use")
```


```{r}
model1 <- lm(electoral_democracy ~ internet_use, data)
summary(model1)
```
```{r message=FALSE, warning=FALSE}
model2 <- lm(gdp ~ internet_use, data)
summary(model2)
```

Creating a table of the results of the regression analysis using `texreg`.
For the first time, install the pacage `texreg`.

```{r eval=FALSE}
install.packages("texreg")
```


```{r}
library(texreg)
```


```{r}
models <- list("Model 1" = model1,
               "Model 2" = model2)
screenreg(models, stars = NULL)
```
