# UN Data {#undata}

UNdata: https://data.un.org

Data of World Development Indicators are in a uniform format and downloadable using an R package WDI. So it is easy to handle. However, other data require data transformation to make it tidy. We give a couple of examples. Most of the UN data, they are in CSV, and you can get a link quickly, or download it by clicking. Though the data structure is not uniform, it is relatively easy to handle.

## Setup

```{r}
library(tidyverse)
```
```{r}
wdi_cache <- read_rds("./data/wdi_cache.RData")
```

## Importing UN Data

Since most of the UN data are available in CSV format, if you obtain the link, i.e., `url`, to the file, you can use `read_csv(url)` to load the data. Alternatively, you can download the file to your download folder. Then move the file to the data directory of your project. You can load it by `read_csv("./data/file_name")`.

## Examples

### Education

By the following, you can see that the first row is not the column name. R gives column names such as ...1, ...2, etc., when the column name is void.

You can copy the link (url) by right click or ctrl+click.

```{r}
url_un_edu <- "https://data.un.org/_Docs/SYB/CSV/SYB65_309_202209_Education.csv"
```

```{r eval=FALSE}
un_edu <- read_csv(url_un_edu)
un_edu %>% slice(1:10)
```
```{r echo=FALSE, eval=FALSE}
un_edu <- read_csv(url_un_edu)
write_csv(un_edu, "./data/un_edu0.csv")
```

```{r echo=FALSE, eval=FALSE}
un_edu <- read_csv("./data/un_edu0.csv")
un_edu %>% slice(1:10)
```

Let is skip the first row by adding `skip = 1`.

```{r eval=FALSE}
un_edu <- read_csv(url_un_edu, skip = 1)
un_edu %>% slice(1:10)
```
```{r echo=FALSE, eval=FALSE}
un_edu <- read_csv(url_un_edu, skip = 1)
write_csv(un_edu, "./data/un_edu.csv")
```

```{r eval=TRUE, echo=FALSE}
un_edu <- read_csv("./data/un_edu.csv")
un_edu %>% slice(1:10)
```

It is a very large data, and we need to check the values.

```{r}
summary(un_edu)
```
We can see that the Year is from 2000 to 2021. The first variable, Region/Country/Area and the fifth variable, Value are `dbl`, i.e., double; hence, these are numerical variables, and you can see them from the summary as well. But it is not easy to see other variables. Let us try them one by one.

```{r}
un_edu %>% distinct(...2)
```

```{r}
un_edu %>% distinct(Series)
```
```{r}
un_edu %>% distinct(Footnotes)
```

```{r}
un_edu %>% distinct(Source)
```

* We may need to handle Footnotes carefully later, but for the first exploration, we need, ...2 for Region, Year, Series, Values.

```{r}
df_un_edu <- un_edu %>%
  select(Region = ...2, Year, Series, Value)
df_un_edu
```

Is there a way to separate regions from countries?

```{r}
df_un_edu %>% left_join(wdi_cache$country, by = c("Region"="country")) %>%
  filter(!is.na(iso2c)) %>% distinct(Region)
```

```{r}
df_un_edu %>% left_join(wdi_cache$country, by = c("Region"="country")) %>%
  filter(is.na(iso2c)) %>% distinct(Region)
```

```{r}
df_un_edu %>% left_join(wdi_cache$country, by = c("Region"="country")) %>%
  filter(is.na(iso2c)) %>% distinct(Region) %>% pull()
```

There are some countries iso2c is not properly assigned. From the list above, Probably, the first 12 are areas and the value contains the aggregated value. 

```{r}
area <- df_un_edu %>% distinct(Region) %>% slice(1:12) %>% pull()
area
```
```{r}
un_edu_area <- df_un_edu %>% filter(Region %in% area)
un_edu_region <- df_un_edu %>% filter(!Region %in% area)
```

Now we can start studying the data.

```{r}
un_edu_area %>% 
  filter(Series %in% c("Gross enrollment ratio - Upper secondary level (male)", "Gross enrollment ratio - Upper secondary level (female)")) %>%
  ggplot(aes(Year, Value, color = Region, linetype = Series)) + geom_line()
```

```{r}
un_edu_area %>% 
  filter(Series %in% c("Gross enrollment ratio - Upper secondary level (male)", "Gross enrollment ratio - Upper secondary level (female)")) %>% 
  pivot_wider(names_from = Series, values_from = Value) %>%
  mutate (Ratio = `Gross enrollment ratio - Upper secondary level (female)`/`Gross enrollment ratio - Upper secondary level (male)`) %>%
  ggplot(aes(Year, Ratio, color = Region, linetype = Region)) + geom_line() +
  labs(title = "Upper Secondary Level Education", subtitle = "Ratio = female/male")
```

### Population

Data structure is similar to the previous one. So use `skip=1`, and check the variable s briefly.


```{r eval=FALSE}
url_un_pop = "https://data.un.org/_Docs/SYB/CSV/SYB65_246_202209_Population%20Growth,%20Fertility%20and%20Mortality%20Indicators.csv"
df_un_pop <- read.csv(url_un_pop, skip = 1)
df_un_pop
```

```{r eval=FALSE, echo=FALSE}
url_un_pop = "https://data.un.org/_Docs/SYB/CSV/SYB65_246_202209_Population%20Growth,%20Fertility%20and%20Mortality%20Indicators.csv"
df_un_pop <- read.csv(url_un_pop, skip = 1)
write_csv(df_un_pop, "./data/un_pop.csv")
```

```{r eval=TRUE, echo=FALSE}
df_un_pop <- read_csv("./data/un_pop.csv")
df_un_pop %>% slice(1:10)
```

```{r}
df_un_pop %>% distinct(Source)
```

```{r}
df_un_pop %>% distinct(Footnotes)
```

```{r}
df_un_pop %>% distinct(X)
```

```{r}
df_un_pop %>% distinct(Series)
```

1. Footnotes need to be studied.
2. There are four different sources.
3. X is for the region, the first 30 are areas, and the rest are countries or regions.
4. There are seven series. It may be easier if we assign shorter names for each value.

```{r}
pop_area <- df_un_pop %>% distinct(X) %>% slice(1:30) %>% pull()
pop_area
```

```{r}
un_pop <- df_un_pop %>% select(Region = X, Year, Series, Value)
un_pop
```

Let us change the names of series.

```{r}
un_pop_wide <- un_pop %>% pivot_wider(names_from = Series, values_from = Value)
colnames(un_pop_wide) <- c("Region", "Year", "IncRate", "Fert", "InfDeath", "MatDeath", "LifeExp", "LifeExpM", "LifeExpF")
un_pop_wide
```

```{r}
un_pop_long <- un_pop_wide %>% pivot_longer(cols = -c(1,2), names_to = "Series", values_to = "Value")
un_pop_long                                           
```


```{r}
un_pop_long_area <- un_pop_long %>% filter(Region %in% pop_area)
un_pop_long_region <- un_pop_long %>% filter(!Region %in% pop_area)
un_pop_wide_area <- un_pop_wide %>% filter(Region %in% pop_area)
un_pop_wide_region <- un_pop_wide %>% filter(!Region %in% pop_area)
```


Now we can visualize data.
