# World Inequality Report {#wir}

## Set up

```{r}
library(tidyverse)
library(readxl) # for excel files
library(WDI)
```

### World Inequility Report - WIR2022

* World Inequality Report: https://wir2022.wid.world/
* Executive Summary: https://wir2022.wid.world/executive-summary/
* Methodology: https://wir2022.wid.world/methodology/
* URL of Executive Summary Data: https://wir2022.wid.world/www-site/uploads/2022/03/WIR2022TablesFigures-Summary.xlsx

Please add `mode="wb"` (web binary). This should work better. 

```{r summary-data, cache=TRUE, eval = FALSE}
url_summary <- "https://wir2022.wid.world/www-site/uploads/2022/03/WIR2022TablesFigures-Summary.xlsx"
download.file(url = url_summary, 
              destfile = "./data/WIR2022s.xlsx", 
              mode = "wb") 
```

If you get an error, download the file directory from the methodology site into your computer, then open it with Excel and save it in the data folder of your R Studio project. Then R studio can recognize it easily as an Excel data.

Generally, a text file such as a CSV file is easy to import, but a binary file is difficult to handle. It is because unless R can recognize its file type, for example, Excel or so, R cannot import the data.

```{r}
excel_sheets("./data/WIR2022s.xlsx")
```


## WIR Package

In the following, we explain how to download data by an R package `wir`. First, you need to install the package. However, it is not an official R package yet; you need to use the package `devtools` to install it.

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("WIDworld/wid-r-tool")
```


I have not studied fully, but you can download the data by a package called `wir`. See [here](https://ds-sl.github.io/data-analysis/wir2022.nb.html). After installing the package, check the [codebook](https://wid.world/codes-dictionary) of the indicators. The following is not the ratio given in F8, but an example. 

* w	wealth-to-income ratio or labor/capital share	fraction of national income
* wealg: net public wealth to net national income ratio
* wealp: net private wealth to net national income ratio

```{r eval=FALSE}
library(wid)
wwealg <- download_wid(indicators = "wwealg", areas = "all", years = "all")
wwealp <- download_wid(indicators = "wwealp", areas = "all", years = "all")
```

```{r eval=FALSE, echo=FALSE}
library(wid)
wwealg <- download_wid(indicators = "wwealg", areas = "all", years = "all")
wwealp <- download_wid(indicators = "wwealp", areas = "all", years = "all")
write_csv(wwealg, "./data/wwealg.csv")
write_csv(wwealp, "./data/wwealp.csv")
```
```{r eval=TRUE, echo=FALSE}
wwealg <-read_csv("./data/wwealg.csv")
wwealp <-read_csv("./data/wwealp.csv")
```

```{r}
public <- wwealg %>% select(country, year, public = value)
public
```

```{r}
private <- wwealp %>% select(country, year, private = value)
private
```

```{r}
public_vs_private <- public %>% left_join(private)
public_vs_private
```

We use `wdi_cache` created by `wdi_cache = WDI::wdi_cache()`.

```{r}
wdi_cache <- read_rds("./data/wdi_cache.RData")
```


```{r}
df_pub_priv <- public_vs_private %>% pivot_longer(cols = c(3,4), names_to = "category", values_to = "value") %>% left_join(wdi_cache$country, by = c("country"="iso2c")) %>%
  select(country = country.y, iso2c = country, year, category, value, region, income, lending)
df_pub_priv
```

```{r}
unique(df_pub_priv$country)
```


```{r}
df_pub_priv %>% 
  filter(country %in% c("Japan", "Norway", "Sweden", "Denmark", "Finland"), year %in% 1970:2020) %>%
  ggplot(aes(year, value, color = country, linetype = category)) + geom_line()
```

We choose two indicators: 'wealg' and 'wealp'. WIR2022 indicators consists of 6 characters; 1 letter code plus 5 letter code. You can find the list in the codebook.

If you want to study WIR2022, please study the [report](https://wir2022.wid.world/), the [codebook](https://wid.world/codes-dictionary), and wir [vignette](https://ds-sl.github.io/data-analysis/wir_download.pdf) together with the [R Notebook](https://ds-sl.github.io/data-analysis/wir2022.nb.html). 

As I mentioned earlier, the data tables used in the report are available from the following page. 

* Methodology: https://wir2022.wid.world/methodology/

